---
title: "4816_pso_prep2"
output: html_document
date: "2026-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Libraries*
```{r}
library(tidyverse)          #tidyverse_2.0.0
library(phyloseq)           #phyloseq_1.52.0  
library(dplyr)              #dplyr_1.1.4
library(ggplot2)            #ggplot2_4.0.0 
library(vegan)              #vegan_2.7-1

library(ggpubr)             #ggpubr_0.6.1
library(rstatix)            #rstatix_0.7.2 
library(multcompView)       #multcompView_0.1-10
library(pairwiseAdonis)     #pairwiseAdonis_0.4.1
library(car)                #car_3.1-3 

library(ggrepel)            #ggrepel_0.9.6
```

*PSO clean up functions*
```{r}
extract_sample_type <- function(id) {
  id_no4 <- sub("^4", "", id) # remove leading '4' if present
  
  if (grepl("^CC", id_no4)) {
    return("corncob")
  } else if (grepl("^WC", id_no4)) {
    return("woodchip")
  } else if (grepl("^INF", id_no4)) {
    return("influent")
  } else {
    return(NA)
  }
}

extract_bioreactor_id <- function(id) {
  # if INF sample, reactor ID is "0"
  id_no4 <- sub("^4", "", id)
  
  if (grepl("^INF", id_no4)) {
    return("0")
  }
  
  # Match Bioreactor ID assuming format: CC_<bioreactorID>_<day> or WC_<bioreactorID>_<day>
  matches <- regmatches(id_no4, regexec("^[A-Z]+_([0-9]+)_", id_no4))
  
  if (length(matches[[1]]) > 1) {
    return(matches[[1]][2])
  } else {
    return(NA)
  }
}

extract_HRT <- function(id) {
  # 1. If starts with "4" → HRT = 4
  if (grepl("^4", id)) {
    return(4)
  }
  
  if (grepl("^INF", id)) {
    num <- as.numeric(sub("^INF(\\d+)$", "\\1", id))
    if (!is.na(num)) {
      if (num <= 20) return(8)
      if (num >= 21) return(16)
    }
    return(NA)
  }
  
  # 2. Extract last number before optional 'W'
  day_str <- sub(".*_(\\d+)W?$", "\\1", id)  # pull numeric from end
  day <- as.numeric(day_str)
  
  # 3. Apply rules for 8 vs 16 HRT
  if (!is.na(day)) {
    if (day >= 6 & day <= 20) {
      return(8)
    } else if (day >= 21) {
      return(16)
    }
  }
  
  return(NA) # if it doesn't match pattern
}

extract_week <- function(id) {
  # 1. If sample starts with "INF", extract Week number after "INF"
  if (grepl("^INF", id)) {
    Week_str <- sub("^INF(\\d+)W?$", "\\1", id)
    Week <- as.numeric(Week_str)
    return(Week)
  }
  # 2. For other sample IDs, extract last number (before optional W at the end)
  Week_str <- sub(".*_(\\d+)W?$", "\\1", id)
  Week <- as.numeric(Week_str)
  if (!is.na(Week)) {
    return(Week)
  }
  return(NA) # return NA if no match
}
```

*PSO read-in & prep* - phyloseq objects were created from '4_pso_prep & 816_pso_prep' Rmds for respective HRTs & sequencing data runs
```{r}
#--------------------------
# 4 HRT read-in & prep
#--------------------------

ps16S_4 <- readRDS("../data/microbial/TV_16S_4_pso.rds")

# adding HRT
sample_data(ps16S_4)$HRT <- 4
ps16S_4_sample_df <- as.data.frame(sample_data(ps16S_4))

# renaming SampleMedia to SampleSource
names(ps16S_4_sample_df)[names(ps16S_4_sample_df) == "SampleMedia"] <- "SampleSource"

# adjust data types
ps16S_4_sample_df$HRT <- as.factor(ps16S_4_sample_df$HRT)
ps16S_4_sample_df$Bioreactor <- as.factor(ps16S_4_sample_df$Bioreactor)
ps16S_4_sample_df$SampleType <- as.factor(ps16S_4_sample_df$SampleType)
ps16S_4_sample_df$Day <- as.numeric(as.character(ps16S_4_sample_df$Day))

sample_data(ps16S_4) <- sample_data(ps16S_4_sample_df)

#--------------------------
# 8 & 16 HRT read-in & prep
#--------------------------
######################################################################################################## <- start of sample ID changes

ps16S_816 <- readRDS("../data/microbial/TV_16S_8_16_pso.rds")

# renaming sample IDs to correspond with actual week - matched ID to date collected through DNA_extraction.xlsx on CyBox
ps16S_816_sample_df <- as.data.frame(sample_data(ps16S_816), stringsAsFactors = FALSE)

sampleID_map <- data.frame(
  old = c(
    "CC_1_2", "CC_2_2", "CC_3_2",
    "CC_1_3",	"CC_2_3", "CC_3_3",
    "CC_1_4",	"CC_2_4", "CC_3_4",
    "WC_4_2",	"WC_5_2", "WC_6_2",
    "WC_4_3",	"WC_5_3", "WC_6_3",
    "WC_4_4", "WC_5_4", "WC_6_4",
    "INF23W"
  ),
  new = c(
    "CC_1_6", "CC_2_6", "CC_3_6",
    "CC_1_10", "CC_2_10", "CC_3_10",
    "CC_1_15", "CC_2_15", "CC_3_15",
    "WC_4_6",	"WC_5_6", "WC_6_6",
    "WC_4_10", "WC_5_10", "WC_6_10", 
    "WC_4_15", "WC_5_15", "WC_6_15",
    "INF23")
)

# create vector with old names that need to be replaced and the replacement sample IDs
lookup <- setNames(sampleID_map$new, sampleID_map$old)

# original pso 8_16 sample IDs
current <- sample_names(ps16S_816)

# replace only where match exists
sample_names(ps16S_816) <- ifelse(current %in% names(lookup),
                                  lookup[current],
                                  current)

# check sample df
ps16S_816_sample_df <- as.data.frame(sample_data(ps16S_816), stringsAsFactors = FALSE)

# rownames to SampleID column
ps16S_816_sample_df$SampleID <- rownames(ps16S_816_sample_df)

# extract new weeks
ps16S_816_sample_df$Week <- sapply(ps16S_816_sample_df$SampleID, extract_week)

# adding back to pso
sample_data(ps16S_816) <- sample_data(ps16S_816_sample_df)

# check sample df
ps16S_816_sample_df <- as.data.frame(sample_data(ps16S_816))


######################################################################################################## <- end of sample ID changes

# adding HRT
sample_data(ps16S_816)$HRT <- ifelse(sample_data(ps16S_816)$Week <= 20, 8, 16)
ps16S_816_sample_df <- as.data.frame(sample_data(ps16S_816))

# renaming "SampleMedia" to "SampleSource"
names(ps16S_816_sample_df)[names(ps16S_816_sample_df) == "SampleMedia"] <- "SampleSource"

# adjust data types
ps16S_816_sample_df$HRT <- as.factor(ps16S_816_sample_df$HRT)
ps16S_816_sample_df$Bioreactor <- as.factor(ps16S_816_sample_df$Bioreactor)
ps16S_816_sample_df$SampleType <- as.factor(ps16S_816_sample_df$SampleType)
ps16S_816_sample_df$Week <- as.numeric(as.character(ps16S_816_sample_df$Week))

sample_data(ps16S_816) <- sample_data(ps16S_816_sample_df)

# check sample df
ps16S_816_sample_df <- as.data.frame(sample_data(ps16S_816))
```

*Merging HRT psos*
```{r}
ps16S_all <- merge_phyloseq(ps16S_4, ps16S_816)

# sample data frames
ps16_sample_df <- as.data.frame(sample_data(ps16S_all))

# saving file
#saveRDS(ps16S_all, file = "../data/microbial/TV_16S_4816_pso.rds")
```

*Original Sample Size df*
```{r}
sample_size_df <- ps16_sample_df |>
  group_by(SampleType, HRT, SampleSource) |>
  summarise(n_samples = n(), .groups = "drop")

print(sample_size_df)
```

*Look at Good, Clean Data - filtering low number of reads*
```{r}
sampling_coverage <- as.data.frame(sample_sums(otu_table(ps16S_all)))
sampling_coverage <- sampling_coverage |>
  rename("sample_sums(otu_table(ps16S_all))" = "n_seqs") |>
  tibble::rownames_to_column("SampleID")

# sampling_coverage <- sampling_coverage |>
#   rename("n_seqs" = "sample_sums(otu_table(ps16S_all))") |>
#   tibble::rownames_to_column("SampleID")

# any natural breaks in the data?
sampling_coverage |>
  ggplot(aes(x = n_seqs)) +
  geom_histogram() +
  coord_cartesian(xlim = c(0, 8000)) +
  ggtitle("Distribution of Reads per Sample [zoomed]") +
  labs(x = "Total Reads", y = "Number of Samples")

# examine outliers (threshold of ~2000)
sampling_coverage |>
  ggplot(aes(x = 1, y = n_seqs)) +
  geom_jitter() +
  scale_y_log10()

# arranging samples by abundance (threshold of ~2000)
sampling_coverage |>
  arrange(n_seqs) |>
  ggplot(aes(x = 1:nrow(sampling_coverage), y = n_seqs)) +
  geom_line() +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 5000))

# filter out by threshold of 2000
sampling_filt <- sampling_coverage |>
  filter(n_seqs > 2000)

samples_to_keep <- sampling_filt$SampleID

ps16S_filt <- subset_samples(ps16S_all, SampleID %in% samples_to_keep) # 304/357 samples kept
ps16S_filt_sample_df <- as.data.frame(sample_data(ps16S_filt))
```
```{r}
# saving filtered file
#saveRDS(ps16S_filt, file = "../data/microbial/TV_16S_4816_filt_pso.rds")
```

*######################################################################*
START OF 301 SAMPLE RDS FILTERING & SAVING (114 MEDIA SAMPLES RESULT)
*######################################################################*

*CC Week 25 Sample Removal*
```{r}
#-------------------------------------------------
# PSO w/ all HRTs REMOVING Week 21 & 25 CC samples
#-------------------------------------------------
samples_to_remove <- c("CC_1_25", "CC_2_25", "CC_3_25")

ps16S_final <- prune_samples(!(sample_names(ps16S_filt) %in% samples_to_remove), ps16S_filt) # 2369 taxa, 301 samples
#saveRDS(ps16S_final, file = "../data/microbial/ps16S_final_301_samps.rds")

# sample data frame
ps16S_final_df <- as.data.frame(sample_data(ps16S_final))
```

*Filtered Sample Size Table*
```{r}
sample_size_filt_df <- ps16S_final_df |>
  group_by(SampleType, HRT, SampleSource) |>
  summarise(n_samples = n(), .groups = "drop")

print(sample_size_filt_df)
```

*Normal & Equal Variance Stat Tests*
```{r}
ps16S_final <- readRDS("../data/microbial/ps16S_final_301_samps.rds")

ps16_sample_df <- as.data.frame(sample_data(ps16S_final))

# extract abundance matrix from phyloseq object
abundance_matrix <- as(otu_table(ps16S_final), "matrix")

# identify columns with zero variance (all identical values)
zero_var_cols <- which(apply(abundance_matrix, 2, var) == 0)

# remove zero-variance taxa
if (length(zero_var_cols) > 0) {
  abundance_matrix_filtered <- abundance_matrix[, -zero_var_cols]
} else {
  abundance_matrix_filtered <- abundance_matrix
}

# generate Q-Q plots for visual inspection
par(mfrow = c(1, 3))
for (i in 1:3) {  # Example for first 3 taxa
  qqnorm(abundance_matrix_filtered[, i], main = colnames(abundance_matrix_filtered)[i])
  qqline(abundance_matrix_filtered[, i])
}

plot(density(abundance_matrix_filtered[, 1]), main = "Taxon Abundance Distribution")

# perform Shapiro-Wilk normality tests - NON NORMAL
shapiro_results <- apply(abundance_matrix_filtered, 2, shapiro.test)
lapply(shapiro_results, function(x) x$p.value)  # p < 0.05 indicates non-normality

# Shapiro-Wilk p-value df
sw_p_values <- sapply(shapiro_results, function(x) x$p.value)
sw_asv_names <- colnames(abundance_matrix_filtered)
shapiro_df <- data.frame(
  ASV = sw_asv_names,
  p_value = sw_p_values,
  row.names = NULL
)

# ------------------------------------------------------------------------------------------------

# perform levene's test for equal variance 
abundance_df <- as.data.frame(abundance_matrix)
abundance_df$Sample <- rownames(abundance_matrix)
abundance_df <- abundance_df |> dplyr::select(Sample, everything())

abundance_long <- pivot_longer(
  abundance_df,
  cols = -Sample,
  names_to = "ASV",
  values_to = "Abundance"
)

abundance_long$SampleType <- sapply(abundance_long$Sample, extract_sample_type)
abundance_long$Bioreactor <- sapply(abundance_long$Sample, extract_bioreactor_id)
abundance_long <- abundance_long |> dplyr::select(Sample, SampleType, Bioreactor, everything())

leveneTest(Abundance ~ Bioreactor, data = abundance_long) 
# p < 0.05 indicates non-equal variance (p = 2.121e-08) <- 301 samples
```