---
title: "4_pso_prep"
output: html_document
date: "2025-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Library*
```{r}
library(BiocManager)      #BiocManager_1.30.26
library(tidyverse)        #tidyverse_2.0.0 
library(phyloseq)         #phyloseq_1.52.0
library(Biostrings)       #Biostrings_2.76.0
library(dada2)            #dada2_1.36.0 
library(ShortRead)        #ShortRead_1.66.0 
```
```
BiocManager::install(version = "3.21")
BiocManager::install("ShortRead")
BiocManager::install("dada2")
```

*Reading in all sample metadata & filtering*
```{r}
# read in all samples
sample_metadata <- read_table("../data/microbial/pso_prep/250806_Howe_16S_250801_ED&MR.txt") 

# filter for TV
sample_metadata <- sample_metadata |>
  filter(grepl("^TV_4", SampleID))

# replace - with _
sample_metadata$SampleID <- gsub("-", "_", sample_metadata$SampleID)
```
```
TV_samples <- TV_samples %>%
  mutate(sampleType = case_when(
    Plate == 1 ~ "Water",
    Plate == 2 ~ "Media",
    TRUE ~ NA_character_  # assigns NA to other cases if needed
  ))
```

*Working Directory & Saving Files*
```{r}
path16S <- "../data/fastq_4HRT_16S"
list.files(path16S) # includes 97 forward & backward seqs (194 total) <-- missing a sample since samplemetadata has 98 rows

## NOTE TO SELF: added a "-" between the "INF" and day number in both the file name and sample metadata txt file manually for INF-0, INF-4, and INF-8
```

*Renaming Files*
```{r}
fnF <- sort(list.files(path16S, pattern = "R1_001.fastq.gz", full.names = TRUE))
fnR <- sort(list.files(path16S, pattern = "R2_001.fastq.gz", full.names = TRUE))

print(fnF) # .fastq.gz files
print(fnR) # .fastq.gz files

# extract sample names, assuming filenames have format: SAMPLENAME_XXX.fastq
sample.names16S <- sapply(strsplit(tools::file_path_sans_ext(basename(fnF)), "_"), function(x) {
  if (grepl("INF", x[2])) {
    # INF case
    return(paste(x[2], x[3], x[4], sep = "_"))
  } else {
    # Non-INF cases
    return(paste(x[2], x[3], x[4], sep = "_"))
  }
})
```
```{r}
print(sample.names16S) # "4INF_11_CCW", "4CC_1_0", "4WC_4_0W"
```

*Inspecting Read Quality Profiles*
```{r, cache=TRUE}
# [1:2] displays the first and second forward read SRR files
plotQualityProfile(fnF[1:3])

# [1:2] displays the first and second reverse read SRR files
plotQualityProfile(fnR[1:3])
```

*Place filtered files in filtered subdirectory*
```{r}
filtFs <- file.path(path16S, "filtered_16S", paste0(sample.names16S, "_F_filt.fastq.gz"))
filtRs <- file.path(path16S, "filtered_16S", paste0(sample.names16S, "_R_filt.fastq.gz"))
names(filtFs) <- sample.names16S
names(filtRs) <- sample.names16S
```

*Naming Checkpoint*
```{r}
print(fnF)
print(fnR)
```
```{r}
print(filtFs)
print(filtRs)
```
```
names(fnF) <- sample.names16S
names(fnR) <- sample.names16S
```

*Filtering Reads* 
```{r}
reads16S <- filterAndTrim(fnF, filtFs, fnR, filtRs, truncLen = c(145,145),
                     maxN = 0, maxEE = c(2,2), truncQ = 2, rm.phix = TRUE, compress = TRUE, multithread = TRUE)

write.csv(reads16S, file = "../data/microbial/pso_prep/4_reads_out_16S.csv")
```

*Error Models (Forward & Reverse)*
```{r}
errF <- learnErrors(filtFs, multithread = TRUE)
# 38316975 total bases in 264255 reads from 97 samples will be used for learning the error rates

errR <- learnErrors(filtRs, multithread = TRUE)
# 38316975 total bases in 264255 reads from 97 samples will be used for learning the error rates
```

*Visualizing estimated error rates*
```{r}
plotErrors(errF, nominalQ = TRUE)

# the estimated error rates (black line) are a good fit to the observed rates (points), and the error rates drop with increased quality as expected. everything looks reasonable and we proceed with confidence.
```

*Apply the core sample inference algorithm to the filtered and trimmed sequence data*
```{r}
# forward
dadaFs <- dada(filtFs, err=errF, multithread=TRUE)
dadaFs[[1]]

# reverse
dadaRs <- dada(filtRs, err=errR, multithread=TRUE)
dadaRs[[1]]
```

*Merge paired reads*
```{r}
# ensure dfs are all same lengths (YES: lengths are 97)
length(filtFs)
length(filtRs)
length(dadaFs)
length(dadaRs)

# if the lengths did not match: run code below (will lose samples)
#filtFs <- filtFs[names(dadaFs)]
#filtRs <- filtRs[names(dadaRs)]

mergers16S <- mergePairs(dadaFs, filtFs, dadaRs, filtRs, verbose=TRUE)
```

*Construct the sequence table out of merged pairs*
```{r}
seqtable16S <- makeSequenceTable(mergers16S)

# returns the dimension (# of columns x # of rows)
dim(seqtable16S)

# inspect distribution of sequence lengths
table(nchar(getSequences(seqtable16S)))

# inspect distribution of sequence lengths
dist_16Sseq_length <- table(nchar(getSequences(seqtable16S)))
barplot(dist_16Sseq_length) 
# since range is only 251 - 255, no additional filtering was done at this stage (highest around 253)
```

*Remove chimeras*
```{r}
seqtable16S.nochim <- removeBimeraDenovo(seqtable16S, method="consensus", multithread=TRUE, verbose=TRUE)

dim(seqtable16S.nochim)
# identified 15 bimeras out of 1100 input sequences.
```

*Track reads through the pipeline*
```{r}
getN <- function(x) sum(getUniques(x))
track16S <- cbind(reads16S, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers16S, getN), rowSums(seqtable16S.nochim))

colnames(track16S) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track16S) <- sample.names16S
head(track16S)
```
```{r}
write.csv(track16S, file = "../data/microbial/pso_prep/TVsample_track_16S_4HRT.csv")
```

*Save seqtable.nochim*
```{r}
saveRDS(seqtable16S.nochim, file = "../data/microbial/pso_prep/4_seqtable_nochim_TV_16S.rds")
```

*Assign taxonomy*
On home computer
- upload "4_seqtable_nochim_TV_16S.rds" onto Nova On Demand (/work/adina/lilliang/DBRLC)


On HPC Nova on Demand R Studio Interactive Session
> library(dada2)

If not working, try: 
> if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("dada2", version = "3.20")
> library(dada2)

After working: Note, must download appropriate taxa database and assign correct pathway
> getwd()  # should be whereever you sent the .rds file to in HPC
> seqtab_nochim <- readRDS("4_seqtable_nochim_TV_16S.rds")
> taxa.ref <- "/work/adina/lilliang/assignTaxa/silva_nr99_v138.1_train_set.fa.gz"
> taxa <- assignTaxonomy(seqtab_nochim, taxa.ref, multithread = TRUE, tryRC = TRUE)
> taxa <- addSpecies(taxa, "/work/adina/lilliang/assignTaxa/silva_species_assignment_v138.1.fa.gz")

multithread = False?????

> saveRDS(taxa, file = "taxa_TV_16S_4HRT.rds")
> quit()


On HPC (outside of R session)
- download "taxa_TV_16S_4HRT.rds" to home computer 

*Read-in "taxa.rds" file*
```{r}
getwd()
file.exists("../data/microbial/pso_prep/taxa_TV_16S_4HRT.rds") #TRUE
taxa16S <- readRDS("../data/microbial/pso_prep/taxa_TV_16S_4HRT.rds")

taxa16S.print <- taxa16S # removing sequence rownames for display only
rownames(taxa16S.print) <- NULL
head(taxa16S.print)
```

*Construct sample dataframe (based on SampleID)*
```{r}
sample_metadata

sample_metadata$SampleID <- sub("^TV_", "", sample_metadata$SampleID)

sample_metadata_full <- sample_metadata |> 
  mutate(
    SampleType = case_when(
      grepl("^4CC", SampleID) ~ "corncob",
      grepl("^4WC", SampleID) ~ "woodchip",
      grepl("^4INF", SampleID) ~ "influent",
      TRUE ~ NA_character_
    ),
    
    Bioreactor = case_when(
      grepl("^4CC_[0-9]+_", SampleID) ~ as.numeric(sub("4CC_([0-9]+)_.*", "\\1", SampleID)),
      grepl("^4WC_[0-9]+_", SampleID) ~ as.numeric(sub("4WC_([0-9]+)_.*", "\\1", SampleID)),
      grepl("^4INF", SampleID) ~ NA_real_,
      TRUE ~ NA_real_
    ),
    
    SampleMedia = case_when(
      grepl("W$", SampleID) ~ "water",
      grepl("^4INF", SampleID) ~ "water",
      TRUE ~ "media"
    ),
    
    Day = as.numeric(case_when(
      grepl("^4INF", SampleID) ~ str_extract(SampleID, "(?<=INF_?)\\d+"),
      
      # WC or CC samples — take last run of digits in string
      grepl("^4?(CC|WC)", SampleID) ~ str_extract(SampleID, "(\\d+)(?=[^0-9]*$)"),
      
      TRUE ~ NA_character_
      )
    )
  )
```

```{r}
#sample name troubleshoot 
## which 1 sample is missing
rownames(sample_metadata_full) <- sample_metadata_full$SampleID

rn1 <- rownames(sample_metadata_full)
rn2 <- rownames(seqtable16S.nochim)

setdiff(rn1, rn2) # 4WC_5_8 is only in sample metadata

# removing missing sample from metadata
sample_table_filt <- sample_metadata_full %>%
  filter(SampleID != "4WC_5_8")

sample_table_filt <- as.data.frame(sample_table_filt)
rownames(sample_table_filt) <- sample_table_filt$SampleID

# rearranging sample ID order of sample filt
sample_table_filt <- sample_table_filt %>%
  arrange(SampleID)

rownames(seqtable16S.nochim)
rownames(sample_table_filt)
```

*Phyloseq object construction*
```{r}
ps16S <- phyloseq(otu_table(seqtable16S.nochim, taxa_are_rows=FALSE), 
               sample_data(sample_table_filt), 
               tax_table(taxa16S))
```

*Assign sequences to "ASV#" ID names*
```{r}
dna16S <- Biostrings::DNAStringSet(taxa_names(ps16S))
names(dna16S) <- taxa_names(ps16S)
ps16S <- merge_phyloseq(ps16S, dna16S)
taxa_names(ps16S) <- paste0("ASV", seq(ntaxa(ps16S)))
ps16S
```

*Save ps as a .rds*
```{r}
saveRDS(ps16S, file = "TV_16S_4_pso.rds")
```
