---
title: "figures_stats"
output: html_document
date: "2026-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Libraries & Analysis Workflow*
```{r}
#---------
# GENERAL
#---------
library(tidyverse)        #tidyverse_2.0.0
library(ggplot2)          #ggplot2_4.0.0 
library(ggpubr)           #ggpubr_0.6.1
library(rstatix)          #rstatix_0.7.2  
library(dplyr)            #dplyr_1.1.4 
library(patchwork)        #patchwork_1.3.2 

#---------------------------
### 1) NRE - (Fig1, FigS1)
# All samples, all HRTs
#---------------------------
library(lubridate)        #lubridate_1.9.4
library(car)              #car_3.1-3 
library(FSA)              #FSA_0.10.0
library(pairwiseAdonis)   #pairwiseAdonis_0.4.1

#--------------------------------------
### 2) NMDS - (Fig2A, FigS2, TableS3)
# Media samples only, all HRTs
#--------------------------------------
library(phyloseq)         #phyloseq_1.52.0
library(vegan)            #vegan_2.7-1
library(webshot2)         #webshot2_0.1.2
library(ggrepel)          #ggrepel_0.9.6 


#--------------------------------------------
### 2.5) Sequence Read Count - writing point
# Media samples only, all HRTs
#--------------------------------------------

#-------------------------------
### 3) VennDiagrams - (Fig2B-D)
# Media samples only, all HRTs
#-------------------------------
library(VennDiagram)      #VennDiagram_1.7.3
library(png)              #png_0.1-8 
library(grid)             #base package

#----------------------------------------------------
# 4) Differential Abundance Analysis - (Fig3A, FigS4)
# Media samples only, 16-h HRT
#----------------------------------------------------
library(ggh4x)            #ggh4x_0.3.1

#-----------------------------------
### 4.5) P/A heatmaps - (Fig3A)
# 317 media samples only, all HRTs
#-----------------------------------
library(ggnewscale)       #ggnewscale_0.5.2

#--------------------------------------------------------
### 4.75) Summed relative abundance of se-ASVs - (FigS5)
# Media samples only, all HRTs
#--------------------------------------------------------
library(ggsignif)         #ggsignif_0.6.4 

#----------------------------------------
### 5) Correlation Analysis - (Fig3B-C)
# Media samples only, all HRTs
#----------------------------------------

#-------------------------------------------
# 6) PICRUSt2 - (Fig4, FigS5)
# 317 se-asvs, media samples only, 16-h HRT
#-------------------------------------------

#-----------------------------------------
### 7) Alpha Diversity - (FigS3, Table S2)
# Media samples only, all HRTs
#-----------------------------------------

#-------------NEW 1/29/2026-------------

#-------------------------------
### 8) TOC
# Media samples only, all HRTs
#-------------------------------

#-------------------------------
### 9) Taxonomy
# Media samples only, all HRTs
#-------------------------------
library(RColorBrewer)     #RColorBrewer_1.1-3 
library(stringr)          #stringr_1.5.2
```

*Functions*
```{r}
# -----------------------
# PSO clean up functions
# -----------------------

extract_sample_type <- function(id) {
  # Remove leading '4' if present
  id_no4 <- sub("^4", "", id)
  
  if (grepl("^CC", id_no4)) {
    return("corncob")
  } else if (grepl("^WC", id_no4)) {
    return("woodchip")
  } else if (grepl("^INF", id_no4)) {
    return("influent")
  } else {
    return(NA)
  }
}

extract_bioreactor_id <- function(id) {
  # If INF sample, reactor ID is "0"
  id_no4 <- sub("^4", "", id)
  
  if (grepl("^INF", id_no4)) {
    return("0")
  }
  
  matches <- regmatches(id_no4, regexec("^[A-Z]+_([0-9]+)_", id_no4))
  
  if (length(matches[[1]]) > 1) {
    return(matches[[1]][2])
  } else {
    return(NA)
  }
}

extract_HRT <- function(id) {
  if (grepl("^4", id)) {
    return(4)
  }
  
  if (grepl("^INF", id)) {
    num <- as.numeric(sub("^INF(\\d+)$", "\\1", id))
    if (!is.na(num)) {
      if (num <= 20) return(8)
      if (num >= 21) return(16)
    }
    return(NA)
  }
  
  day_str <- sub(".*_(\\d+)W?$", "\\1", id)  # pull numeric from end
  day <- as.numeric(day_str)
  
  if (!is.na(day)) {
    if (day >= 6 & day <= 20) {
      return(8)
    } else if (day >= 21) {
      return(16)
    }
  }
  
  return(NA) # if it doesn't match pattern
}

extract_sample_source <- function(id) {
  # Check if sample ID ends with "W"
  if (grepl("W$", id)) {
    return("water")
  }
  if (grepl("^INF", id)) {
    return("water")
  }
  else {
    return("media")
  }
}

extract_week <- function(id) {
  # 1. If sample starts with "INF", extract Week number after "INF"
  if (grepl("^INF", id)) {
    Week_str <- sub("^INF(\\d+)W?$", "\\1", id)
    Week <- as.numeric(Week_str)
    return(Week)
  }
  # 2. For other sample IDs, extract last number (before optional W at the end)
  Week_str <- sub(".*_(\\d+)W?$", "\\1", id)
  Week <- as.numeric(Week_str)
  if (!is.na(Week)) {
    return(Week)
  }
  return(NA) # return NA if no match
}

# -------------
# VennDiagrams
# -------------
# convert ps to presence-absence (1 if present, 0 is not)
to_presence_absence <- function(x) {
  x[x > 0] <- 1
  return(x)
}

# ------------------------------
# PICRUSt2 Downstream Functions
# ------------------------------

# function to convert sample columns to rows
function_to_long <- function(df) {
  df |>
    rename(`function` = 1) |>
    pivot_longer(-`function`, names_to = "SampleID", values_to = "Abundance")
}


# calculate relative abundance within each sample
relative_abundance <- function(df_long) {
  df_long |>
    group_by(SampleID) |>
    mutate(RelAbund = Abundance / sum(Abundance, na.rm = TRUE)) |>
    ungroup()
}
```

*################################################################################################################*
*################################################################################################################*
#---------------------------
### 1) NRE - (Fig1, FigS1)
# All samples, all HRTs
#---------------------------

*NRE & GHG Prep*
```{r}
#---------------------------------------------------------
# NRE data read in, string type changes, factor ordering
#---------------------------------------------------------
nre_data <- read_csv("../data/performance/full_NRE_237_data_DBRLC_2026.csv")

# converting columns to numeric/factor
columns_to_num <- c("NOx_conc", "NO2_conc", "NO3_conc", "Ammonia_ammonia_conc", 
                        "DRP_conc", "TOC_conc", "pH", "DO_conc")
columns_to_factor <- c("sample_type", "bioreactor_ID", "influent_id", "HRT")

nre_data <- nre_data |>
  mutate_at(vars(all_of(columns_to_num)), ~ as.numeric(as.character(.)))

nre_data <- nre_data |>
  mutate_at(vars(all_of(columns_to_factor)), ~ as.factor(as.character(.)))

# ordering HRT
HRT_order <- c("4", "8", "16")

nre_data <- nre_data |>
  mutate(HRT = factor(HRT, levels = HRT_order))

#---------------------------------------------------------
# GHG data read in, string type changes, factor ordering
#---------------------------------------------------------
ghg_data <- read_csv("../data/ghg/GHG_data_checked.csv")

# converting columns to factor
columns_to_factor <- c("HRT", "SampleType", "pore_vol", "Time", "Bioreactor")

ghg_data <- ghg_data |>
  mutate_at(vars(all_of(columns_to_factor)), ~ as.factor(as.character(.)))

# ordering HRT
HRT_order <- c("4", "8", "16")

ghg_data <- ghg_data |>
  mutate(HRT = factor(HRT, levels = HRT_order))
```

*Figure 1A & Stats*
```{r, fig.width=6, fig.height=6}
#------------------------------------------------
# Plot - NRE comparison by bioreactor rep (1-6)
#------------------------------------------------
fig1A <- nre_data |>
  ggplot(aes(bioreactor_ID, removal_efficiency, fill = sample_type)) +
  geom_boxplot() +
  facet_wrap(~HRT) +
  labs(
    x = "Bioreactor ID",
    y = expression("NO"[3]^"  -"~removal~efficiency~" (%)")
  ) +
  theme_classic() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        panel.grid = element_blank(),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

#------------------------------------
# Presentation Plot - same as fig 1A
#------------------------------------
pfig1A <- nre_data |>
  ggplot(aes(bioreactor_ID, removal_efficiency, fill = sample_type)) +
  geom_boxplot() +
  facet_wrap(~HRT) +
  labs(
    x = "Bioreactor ID",
    y = expression("NO"[3]^"  -"~removal~efficiency~" (%)")
  ) +
  theme_bw() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        panel.grid = element_blank(),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

#ggsave("../present_figs/fig1a_NRE.png", plot = pfig1A, width = 8, height = 6, units = "in", dpi = 300)
#ggsave("../present_figs/fig1a_6x6_NRE.png", plot = pfig1A, width = 6, height = 6, units = "in", dpi = 300)

# numeric means - NRE & TOC by sample type & HRT
nre_data |>
  group_by(sample_type, HRT) |>
  summarise(mean_removal_efficiency = mean(removal_efficiency, na.rm = TRUE),
            nre_std = sd(removal_efficiency, na.rm = TRUE),
            mean_conc= mean(TOC_conc, na.rm = TRUE),
            toc_std = sd(TOC_conc, na.rm = TRUE))
```
```{r}
#------------------------------------
# Stats - NRE separated by CC vs WC
#------------------------------------

# full model comparison [KRUSKAL] - p < 0.05 = signif diff in removal efficiency between sample type
nre_data |>
  group_by(HRT) |>
  kruskal_test(removal_efficiency ~ sample_type)
```

*Figure 1B & Stats*
```{r, fig.width=4, fig.height=6}
#----------------------------------------------------------
# Plot - NRE comparison across HRT (for each bioreactor)
#----------------------------------------------------------
fig1B <- nre_data |>
  ggplot(aes(HRT, removal_efficiency)) +
  geom_boxplot(aes(fill = sample_type)) +
  labs(
    x = "HRT (hr)",
    y = expression("NO"[3]^"  -"~removal~efficiency~" (%)")
  ) +
  facet_wrap(~bioreactor_ID) +
  theme_classic() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

#------------------------------------
# Presentation Plot - same as fig 1B
#------------------------------------
pfig1B <- nre_data |>
  ggplot(aes(HRT, removal_efficiency)) +
  geom_boxplot(aes(fill = sample_type)) +
  labs(
    x = "HRT (hr)",
    y = expression("NO"[3]^"  -"~removal~efficiency~" (%)")
  ) +
  facet_wrap(~bioreactor_ID) +
  theme_bw() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        panel.grid = element_blank(),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

#ggsave("../present_figs/fig1b_NRE.png", plot = pfig1B, width = 6, height = 6, units = "in", dpi = 300)


nre_data |>
  ggplot(aes(HRT, removal_efficiency)) +
  geom_boxplot(aes(fill = sample_type)) +
  labs(
    x = "HRT (hr)",
    y = expression("NO"[3]^"  -"~removal~efficiency~" (%)")
  ) +
  theme_bw() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        panel.grid = element_blank(),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))
```
```{r}
#----------------------------------------------------------
# Stats - NRE comparison across HRT (for each bioreactor)
#----------------------------------------------------------

# full model comparison [KRUSKAL] - p < 0.05 = signif diff in removal efficiency between HRT
nre_data |>
  group_by(bioreactor_ID) |>
  kruskal_test(removal_efficiency ~ HRT) 

# pairwise comparison [DUNN] 
nre_dunn_br <- nre_data |>
  group_by(bioreactor_ID) |>
  dunn_test(removal_efficiency ~ HRT, p.adjust.method = "bonferroni")

nre_dunn_br_sig <- nre_dunn_br |>
  filter(p.adj.signif != "ns")
```

*FINAL MAIN: Figure 1*
```{r, fig.width=10, fig.height=9}}
fig1 <- fig1A + plot_spacer() + fig1B +
  plot_layout(widths = c(2, 0.1, 1.3)) 

# save plot
#ggsave("../raw_figs_301/dnbr_fig1_NRE.png", plot = fig1, width = 10, height = 6, units = "in", dpi = 300)
```

*Figure S1 A-C & Stats*
```{r, fig.width=4, fig.height=6}
#------------------------------------------------
# Plot - NO2 comparison by bioreactor rep (1-6)
#------------------------------------------------
sfig1a <- nre_data |>
  ggplot(aes(bioreactor_ID, NO2_conc, fill = sample_type)) +
  geom_boxplot() +
  facet_wrap(~HRT) +
  labs(
#    title = "NO2 conc in CC and WC bioreactors",
    x = "Bioreactor ID",
    y = expression("NO"[2]^"  -"~conc~" (mg N/L)")
  ) +
  theme_classic() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

#--------------------------------------------------
# Plot - N2O GHG comparison by bioreactor ID (1-6)
#--------------------------------------------------
sfig1b <- ghg_data |>
  ggplot(aes(Bioreactor, g_n2o, fill = SampleType)) +
  geom_boxplot() +
  facet_wrap(~HRT) +
  labs(
#    title = "N2O emmissions from CC and WC bioreactor replicates", 
    x = "Media Type", 
    y = expression("dissolved N"[2]~"O"~(CO[2]~"e")~("g m"^" -3"~"day"^" -1"))
  ) +
#  scale_y_log10() +
  theme_classic() +
  scale_fill_manual(name = "Media Type",
                     values = c("#ffa52d", "darkseagreen")) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))
        
#------------------------------------------------------
# Plot - Ammonia conc comparison by bioreactor ID (1-6)
#------------------------------------------------------
sfig1c <- nre_data |>
  ggplot(aes(bioreactor_ID, Ammonia_ammonia_conc, fill = sample_type)) +
  geom_boxplot() +
  facet_wrap(~HRT) +
  labs(
#    title = "Ammonia/Ammonia conc in CC and WC bioreactors",
    x = "Bioreactor ID",
    y = "Ammonia/Ammonium Conc (mg N/L)"
  ) +
  theme_classic() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))
```
```{r}
#------------------
# Stats - N2O GHG 
#-----------------
# numeric means
n20_means_df <- ghg_data |>
  group_by(SampleType, HRT) |>
  summarise(mean_conc= mean(g_n2o, na.rm = TRUE),
            std = sd(g_n2o, na.rm = TRUE))

# full model comparison [KRUSKAL] - p < 0.05 = signif diff in removal efficiency between sample type [Only @8h HRT is diff signif]
ghg_data |>
  group_by(HRT) |>
  kruskal_test(g_n2o ~ SampleType)

# pairwise comparison [DUNN] 
ghg_dunn <- ghg_data |>
  group_by(HRT) |>
  dunn_test(g_n2o ~ Bioreactor, p.adjust.method = "bonferroni")

ghg_dunn_sig <- ghg_dunn |>
  filter(p.adj.signif != "ns")

#-------------
# Stats - NO2 
#-------------
# numeric means
no2_means_df <- nre_data |>
  group_by(sample_type, HRT) |>
  summarise(mean_conc= mean(NO2_conc, na.rm = TRUE),
            std = sd(NO2_conc, na.rm = TRUE))

# full model comparison [KRUSKAL] - p < 0.05 = signif diff in removal efficiency between sample type [Only @4h HRT is diff signif]
nre_data |>
  group_by(HRT) |>
  kruskal_test(NO2_conc ~ sample_type) 

# pairwise comparison [DUNN] 
no2_dunn <- nre_data |>
  group_by(HRT) |>
  dunn_test(NO2_conc ~ bioreactor_ID, p.adjust.method = "bonferroni")

no2_dunn_sig <- no2_dunn |>
  filter(p.adj.signif != "ns")

#-----------------
# Stats - Ammonia 
#-----------------

# numeric means
nh3_means_df <- nre_data |>
  group_by(sample_type, HRT) |>
  summarise(mean_conc= mean(Ammonia_ammonia_conc, na.rm = TRUE),
            std = sd(NO2_conc, na.rm = TRUE))

# full model comparison [KRUSKAL] - p < 0.05 = signif diff in removal efficiency between sample type [Only @16h HRT is diff signif]
nre_data |>
  group_by(HRT) |>
  kruskal_test(Ammonia_ammonia_conc ~ sample_type) 

# pairwise comparison [DUNN] 
nh3_dunn <- nre_data |>
  group_by(HRT) |>
  dunn_test(Ammonia_ammonia_conc ~ bioreactor_ID, p.adjust.method = "bonferroni")

nh3_dunn_sig <- nh3_dunn |>
  filter(p.adj.signif != "ns")
```

*FINAL SUPP: Figure S1*
```{r, fig.width=20, fig.height=10}}
figS1 <- sfig1a + plot_spacer() + sfig1b + plot_spacer() + sfig1c +
  plot_layout(widths = c(2, 0.1, 2, 0.1, 2), ncol = 5) 

# save plot
#ggsave("../raw_figs_301/dnbr_sfig1_N_spp.png", plot = figS1, width = 10, height = 6, units = "in", dpi = 300)
```

*################################################################################################################*
*################################################################################################################*
#--------------------------------------
### 2) NMDS - (Fig2A, FigS2, TableS3)
# Media samples only, all HRTs
#--------------------------------------

*PSO read-in*
```{r}
ps16S_final <- readRDS("../data/microbial/ps16S_final_301_samps.rds")
ps16_sample_df <- as.data.frame(sample_data(ps16S_final)) 

# media only
ps_media <- subset_samples(ps16S_final, SampleSource == "media") # 114 samples
ps_media_samp_df <- sample_data(ps_media)
```

*MAIN: Figure 2A*
```{r, fig.width=7, fig.height=6}
#---------------------------
# NMDS prep & calculations 
#---------------------------
bc_dist <- phyloseq::distance(ps_media, method = "bray")
set.seed(321)
nmds_bc <- ordinate(ps_media, method = "NMDS", distance = bc_dist)

nmds_color_map <- c(
  "corncob" = "#ffa52d",
  "woodchip" = "darkseagreen"
)

# stress value
nmds_bc$stress # 0.1524762

#---------------------------------------------------------
# Plot - NMDS (comparison between sample type) 
#---------------------------------------------------------
fig2a <- plot_ordination(ps_media, nmds_bc, type = "samples", color = "SampleType", shape = "SampleSource") +
  geom_point(size = 2) +
  scale_color_manual(values = nmds_color_map, name = "Media Type") +
  scale_shape_manual(values = c("media" = 5), guide = "none") +
  facet_wrap(~ HRT) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 13),
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/dnbr_fig2a_nmds.png", plot = fig2a, width = 7, height = 6, units = "in", dpi = 300)

#------------------------------------
# Presentation Plot - same as fig 2A
#------------------------------------
pfig2a <- plot_ordination(ps_media, nmds_bc, type = "samples", color = "SampleType", shape = "SampleSource") +
  geom_point(size = 2) +
  scale_color_manual(values = nmds_color_map, name = "Media Type") +
  scale_shape_manual(values = c("media" = 5), guide = "none") +
  facet_wrap(~ HRT) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 13),
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../present_figs/fig2a_nmds.png", plot = pfig2a, width = 7, height = 6, units = "in", dpi = 300)
```
```{r}
set.seed(321)
#--------------------------------------
# STATS - NMDS (Within HRT (CC vs WC))
#--------------------------------------
# overall test (adonis2)
metadata <- as(sample_data(ps_media_samp_df), "data.frame")

adonis2(
  as.matrix(bc_dist) ~ SampleType + HRT,
  data = metadata,
  permutations = 999,
  by = "margin"
)

# post hoc test (pairwise adonis) <- TABLE 3A
unique_HRTs <- unique(metadata$HRT)
pairwise_results_sampleType_by_HRT <- list()

for (hrt_level in unique_HRTs) {
  subset_samples <- rownames(metadata)[metadata$HRT == hrt_level]
  if (length(subset_samples) > 1) {  
    bc_subset <- as.matrix(bc_dist)[subset_samples, subset_samples]
    type_subset <- metadata[subset_samples, ]$SampleType
    pw_res <- pairwise.adonis(
      x = bc_subset,
      factors = type_subset,
      p.adjust.m = "BH"
    )
    list_name <- paste0("HRT_", hrt_level)
    pairwise_results_sampleType_by_HRT[[list_name]] <- pw_res
  }
}
```

*Outlier NMDS plots*
```{r, fig.width=7, fig.height=6}
#----------------------------------------------------------------------
# Plot - NMDS (comparison between sample type) <- TOC outliers labeled
#----------------------------------------------------------------------

TOC_outliers <- c("CC_1_28", "CC_2_28", "CC_3_28", "WC_4_7")

# Get ordination coordinates as a data frame
nmds_toc_df <- plot_ordination(ps_media, nmds_bc, type = "samples", 
                          color = "SampleType", shape = "SampleSource")$data

TOC_outliers_df <- nmds_toc_df |> 
  filter(SampleID %in% TOC_outliers) 

TOC_lab_nmds <- plot_ordination(ps_media, nmds_bc, type = "samples", color = "SampleType", shape = "SampleSource") +
  geom_text_repel(aes(label = sample_names(ps_media)), max.overlaps = Inf) +
  geom_point(size = 2) +
  scale_color_manual(values = nmds_color_map, name = "Media Type") +
  scale_shape_manual(values = c("media" = 5), guide = "none") +
  facet_wrap(~ HRT) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 13),
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

toc_nmds_p <- plot_ordination(ps_media, nmds_bc, type = "samples", color = "SampleType", shape = "SampleSource") +
  geom_text_repel(
    data = TOC_outliers_df,
    aes(label = SampleID),  
    max.overlaps = Inf,
    size = 3, 
    nudge_y = -0.2, 
    nudge_x = 0.8) +
  geom_point(size = 2) +
  scale_color_manual(values = nmds_color_map, name = "Media Type") +
  scale_shape_manual(values = c("media" = 5), guide = "none") +
  facet_wrap(~ HRT) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 13),
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15))
```

*Figure S2 A-B & Stats*
```{r, fig.width=7, fig.height=8}
#---------------------------------------------------------
# Plot - NMDS (comparison between sample type/bioreactor) 
#---------------------------------------------------------
sfig2b <- plot_ordination(ps_media, nmds_bc, type = "samples", color = "Bioreactor") +
  geom_point(size = 3) +
  scale_colour_manual(
    values = c(
      "1" = "chocolate", 
      "2" = "orange",
      "3" = "chocolate4",
      "4" = "darkseagreen",
      "5" = "cadetblue",
      "6" = "aquamarine4"),
    name = "Bioreactor ID") +
  facet_wrap(~ HRT) +
  theme_bw() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))

#---------------------------------------------------------
# Plot - NMDS (comparison between sample type & HRT) 
#---------------------------------------------------------
sfig2a <- plot_ordination(ps_media, nmds_bc, type = "samples", color = "SampleType", shape = "HRT") +
  geom_point(size = 3) +
  scale_color_manual(values = nmds_color_map, name = "Media Type") +
  scale_shape_manual(values = c("4" = 5, "8" = 16, "16" = 2)) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 15),
        panel.grid = element_blank(),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))

#ggsave("../present_figs/sfig2a_nmds.png", plot = sfig2a, width = 6, height = 6, units = "in", dpi = 300)
```
```{r}
set.seed(321)
#---------------------------------------------------------
# STATS - NMDS (Within HRT (Bioreactor vs Bioreactor)) 
#---------------------------------------------------------
# overall test (adonis2) 
metadata <- as(sample_data(ps_media_samp_df), "data.frame")

adonis2(
  as.matrix(bc_dist) ~ Bioreactor + HRT,
  data = metadata,
  permutations = 999,
  by = "margin"
)

# post hoc test (pairwise adonis) 
unique_HRTs <- unique(metadata$HRT)
pairwise_results_bioreactors_by_HRT <- list()

for (hrt_level in unique_HRTs) {
  subset_samples <- rownames(metadata)[metadata$HRT == hrt_level]
  if (length(subset_samples) > 1) {  
    bc_subset <- as.matrix(bc_dist)[subset_samples, subset_samples]
    bioreactor_subset <- metadata[subset_samples, ]$Bioreactor
    pw_res <- pairwise.adonis(
      x = bc_subset,
      factors = bioreactor_subset,
      p.adjust.m = "BH"
    )
    list_name <- paste0("HRT_", hrt_level)
    pairwise_results_bioreactors_by_HRT[[list_name]] <- pw_res
  }
}

pairwise_BR_4 <- pairwise_results_bioreactors_by_HRT$HRT_4
pairwise_BR_8 <- pairwise_results_bioreactors_by_HRT$HRT_8
pairwise_BR_16 <- pairwise_results_bioreactors_by_HRT$HRT_16

#---------------------------------------------------------
# STATS - NMDS (Within CC & WC (HRT vs HRT)) 
#---------------------------------------------------------
set.seed(321)
# post hoc test (pairwise adonis) - 
unique_sample_types <- unique(metadata$SampleType)
pairwise_results_HRT_by_sampletype <- list()

for (sample_type in unique_sample_types) {
  subset_samples <- rownames(metadata)[metadata$SampleType == sample_type]
  if (length(subset_samples) > 1) { 
    bc_subset <- as.matrix(bc_dist)[subset_samples, subset_samples]
    hrt_subset <- metadata[subset_samples, ]$HRT
    pw_res <- pairwise.adonis(
      x = bc_subset,
      factors = hrt_subset,
      p.adjust.m = "BH"
    )
    list_name <- paste0("SampleType_", sample_type)
    pairwise_results_HRT_by_sampletype[[list_name]] <- pw_res
  }
}
```

*FINAL SUPP: Figure S2*
```{r, fig.width=16, fig.height=12}
sfig2 <- sfig2a + plot_spacer() + sfig2b +
  plot_layout(ncol = 3, widths = c(1, 0.1, 1))

#ggsave("../raw_figs_301/dnbr_sfig2_nmds.png", plot = sfig2, width = 12, height = 6, units = "in", dpi = 300)
```

*Rarefied Analysis - media only - not saved*
```{r}
# rarefying
min_depth <- min(sample_sums(ps_media))

ps_m_rare <- rarefy_even_depth(ps_media, 
                                 sample.size = min_depth, 
                                 rngseed = 711, 
                                 replace = FALSE) # 895OTUs were removed bc no longer present in any sample after random subsampling

ps_m_rare_df <- sample_data(ps_m_rare)

# calculations
bc_dist_rare <- phyloseq::distance(ps_m_rare, method = "bray")
set.seed(343)
nmds_bc_rare <- ordinate(ps_m_rare, method = "NMDS", distance = bc_dist_rare)

nmds_rare_color_map <- c(
  "corncob" = "#ffa52d",
  "woodchip" = "darkseagreen"
)

nmds_bc_rare$stress # 0.1517114

# plot
plot_ordination(ps_m_rare, nmds_bc_rare, type = "samples", color = "SampleType") +
  geom_point(size = 2) +
  scale_color_manual(values = nmds_rare_color_map, name = "Sample Type") +
  facet_wrap(~ HRT) +
  theme_bw() +
  theme(
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

# stats ----------------------------------------------------------------------------

# overall test (adonis2)
metadata <- as(sample_data(ps_m_rare_df), "data.frame")

adonis2(
  as.matrix(bc_dist_rare) ~ SampleType + HRT,
  data = metadata,
  permutations = 999,
  by = "margin"
)

# post hoc test (pairwise adonis)
unique_HRTs <- unique(metadata$HRT)
pairwise_results_sampleType_by_HRT <- list()

for (hrt_level in unique_HRTs) {
  subset_samples <- rownames(metadata)[metadata$HRT == hrt_level]
  if (length(subset_samples) > 1) {  
    bc_subset <- as.matrix(bc_dist_rare)[subset_samples, subset_samples]
    type_subset <- metadata[subset_samples, ]$SampleType
    pw_res <- pairwise.adonis(
      x = bc_subset,
      factors = type_subset,
      p.adjust.m = "BH"
    )
    list_name <- paste0("HRT_", hrt_level)
    pairwise_results_sampleType_by_HRT[[list_name]] <- pw_res
  }
}
```

*################################################################################################################*
*################################################################################################################*
#--------------------------------------------
### 2.5) Sequence Read Count - writing point
# Media samples only, all HRTs
#--------------------------------------------

*Prep, Plot, & Stats - not saved*
```{r}
# adding total reads as a metadata colummn
sample_data(ps_media)$TotalReads <- sample_sums(ps_media)

# sample data frame
ps_media_samp_df <- data.frame(sample_data(ps_media), stringsAsFactors = FALSE)

# converting cols to factors
columns_to_factor <- c("SampleType", "HRT", "Bioreactor")

ps_media_samp_df <- ps_media_samp_df |>
  mutate_at(vars(all_of(columns_to_factor)), ~ as.factor(as.character(.)))

# ordering HRT
HRT_order <- c("4", "8", "16")

ps_media_samp_df <- ps_media_samp_df |>
  mutate(HRT = factor(HRT, levels = HRT_order))

# summarize total reads by HRT & SampleType
read_summ <- ps_media_samp_df |>
  group_by(HRT, SampleType) |>
  dplyr::summarize(Total_Reads = sum(TotalReads, na.rm=TRUE)) |>
  ungroup()

# taking average
read_avg <- ps_media_samp_df |>
  group_by(HRT, SampleType) |>
  dplyr::summarize(Avg_Reads = mean(TotalReads, na.rm=TRUE)) |>
  ungroup()

# plot (HRT facet) ----------------------------------------------------------------------------
stat_test <- ps_media_samp_df %>%
  group_by(HRT) %>%
  wilcox_test(TotalReads ~ SampleType) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_xy_position(x = "SampleType")

stat_test$symbol <- "ns"

ps_media_samp_df |>
  ggplot(aes(x = SampleType, y = TotalReads, fill = SampleType)) +
  geom_boxplot() +
  labs(x = "SampleType", y = "Average Sequence Read Count", fill="Sample Type"
       ) +
  facet_wrap(~ HRT) +
  theme_classic() +
    scale_fill_manual(values = c(corncob = "#ffa52d", woodchip = "darkseagreen")) +
  stat_pvalue_manual(
    stat_test,
    label = "symbol",
    tip.length = 0.01,
    hide.ns = FALSE 
  ) +
  theme(legend.position = "bottom",
        axis.text.x = element_blank())

# plot (sample type facet) ----------------------------------------------------------------------------
ps_media_samp_df %>%
  group_by(SampleType) %>%
  wilcox_test(TotalReads ~ HRT) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_xy_position(x = "SampleType")

ps_media_samp_df |>
  ggplot(aes(x = HRT, y = TotalReads, fill = SampleType)) +
  geom_boxplot() +
  labs(x = "SampleType", y = "Average Sequence Read Count", fill="Sample Type"
       ) +
  facet_wrap(~ SampleType) +
  theme_classic() +
    scale_fill_manual(values = c(corncob = "#ffa52d", woodchip = "darkseagreen")) +
  theme(legend.position = "bottom",
        axis.text.x = element_blank())
```

*################################################################################################################*
*################################################################################################################*
#-------------------------------
### 3) VennDiagrams - (Fig2B-D)
# Media samples only, all HRTs
#-------------------------------

*FINAL MAIN: Figure 2 B-D*
```{r}
VD_groups <- c("corncob", "woodchip")

VD_media_color_map <- c(
  "CC" = "#ffa52d",
  "WC"= "darkseagreen"
)

#----------------
# 16 HRT - media
#----------------
ps_media_16 <- subset_samples(ps_media, HRT == "16")

# convert to p/a
pa_ps_media_16 <- transform_sample_counts(ps_media_16, to_presence_absence) #otu_table(pa_ps_media_16)

# set of taxa present in any sample in each group
group_taxa_media <- lapply(VD_groups, function(gr) {
  keep <- sample_data(pa_ps_media_16)$SampleType == gr
  subsetted <- prune_samples(keep, pa_ps_media_16)
  presence <- taxa_sums(subsetted) > 0
  taxa_names(pa_ps_media_16)[presence]
})

names(group_taxa_media) <- VD_groups   # labels lists in 'group_taxa'

media_short_labels <- c("CC", "WC")
names(media_short_labels) <- VD_groups

# plot (saves to computer)
venn.diagram(
  x = group_taxa_media, 
  category.names = media_short_labels,
  filename = "../raw_figs_301/dnbr_fig2d_16_VD.png",
  output = "FALSE",
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 350,
  compression = "lzw",

  lwd = 2,
  lty = 'blank',
  fill = VD_media_color_map,

  cex = .5,
  fontface = "bold",
  fontfamily = "sans",

  cat.cex = 0.4,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.fontfamily = "sans"
)

#----------------
# 8 HRT - media
#----------------
ps_media_8 <- subset_samples(ps_media, HRT == "8")

# convert to p/a
pa_ps_media_8 <- transform_sample_counts(ps_media_8, to_presence_absence) #otu_table(pa_ps_media_8)

# set of taxa present in any sample in each group
group_taxa_media <- lapply(VD_groups, function(gr) {
  keep <- sample_data(pa_ps_media_8)$SampleType == gr
  subsetted <- prune_samples(keep, pa_ps_media_8)
  presence <- taxa_sums(subsetted) > 0
  taxa_names(pa_ps_media_8)[presence]
})

names(group_taxa_media) <- VD_groups   # labels lists in 'group_taxa'

media_short_labels <- c("CC", "WC")
names(media_short_labels) <- VD_groups

# plot (saves to computer)
venn.diagram(
  x = group_taxa_media, 
  category.names = media_short_labels,
  filename = "../raw_figs_301/dnbr_fig2c_8_VD.png",
  output = "FALSE",
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 350,
  compression = "lzw",

  lwd = 2,
  lty = 'blank',
  fill = VD_media_color_map,

  cex = .5,
  fontface = "bold",
  fontfamily = "sans",

  cat.cex = 0.4,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.fontfamily = "sans"
)

#----------------
# 4 HRT - media
#----------------
ps_media_4 <- subset_samples(ps_media, HRT == "4")

# convert to p/a
pa_ps_media_4 <- transform_sample_counts(ps_media_4, to_presence_absence) #otu_table(pa_ps_media_4)

# set of taxa present in any sample in each group
group_taxa_media <- lapply(VD_groups, function(gr) {
  keep <- sample_data(pa_ps_media_4)$SampleType == gr
  subsetted <- prune_samples(keep, pa_ps_media_4)
  presence <- taxa_sums(subsetted) > 0
  taxa_names(pa_ps_media_4)[presence]
})

names(group_taxa_media) <- VD_groups   # labels lists in 'group_taxa'

media_short_labels <- c("CC", "WC")
names(media_short_labels) <- VD_groups

# plot (saves to computer)
venn.diagram(
  x = group_taxa_media, 
  category.names = media_short_labels,
  filename = "../raw_figs_301/dnbr_fig2b_4_VD.png",
  output = "FALSE",
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 350,
  compression = "lzw",

  lwd = 2,
  lty = 'blank',
  fill = VD_media_color_map,

  cex = .5,
  fontface = "bold",
  fontfamily = "sans",

  cat.cex = 0.4,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.fontfamily = "sans"
)
```

*################################################################################################################*
*################################################################################################################*
#----------------------------------------------------
# 4) Differential Abundance Analysis - (Fig3A, FigS4)
# Media samples only, 16-h HRT
#----------------------------------------------------

*Prep*
```{r}
ps_media_16 <- subset_samples(ps_media, HRT == "16")

# relative abundance (convert raw counts to relative abundances within each sample)
ps_media_16_rel <- transform_sample_counts(ps_media_16, function(x) x / sum(x))

# melt to long table
ps_media_16_melt <- psmelt(ps_media_16_rel)

# pulling taxonomic info from melted df
taxa_unique <- ps_media_16_melt |>
  distinct(OTU, Kingdom, Phylum, Class, Order, Family, Genus, Species)

# compute group means (CC vs. WC) - (calculate the mean relative abundance of each taxon (OTU) grouped by sample type - one average relative abundance per OTU per SampleType)
mean_abund <- ps_media_16_melt |>
  group_by(OTU, SampleType) |>
  summarise(mean_rel = mean(Abundance), .groups = "drop") |>
  pivot_wider(names_from = SampleType,
              values_from = mean_rel,
              values_fill = 0)

# add back taxnomic columns
mean_abund <- mean_abund |>
  left_join(taxa_unique, by = c("OTU")) |>
  dplyr::select(OTU, corncob, woodchip, Kingdom, Phylum, Class, Order, Family, Genus, Species)

# differential test (Wilcoxon)
pvals <- ps_media_16_melt |>
  group_by(OTU) |>
  summarise(
    cc_vals = list(Abundance[SampleType == "corncob"]),
    wc_vals = list(Abundance[SampleType == "woodchip"]),
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(
    p = if (length(unlist(cc_vals)) > 0 &&
            length(unlist(wc_vals)) > 0) {
      wilcox.test(unlist(cc_vals), unlist(wc_vals))$p.value
    } else {
      NA_real_
    }
  ) |>
  dplyr::select(OTU, p)

mean_abund <- left_join(mean_abund, pvals, by = "OTU")
mean_abund <- mean_abund |>
  mutate(
    sig = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ ""
  ))

# preparing for plot (all sig)
mean_abund_sig <- mean_abund |>
  filter(sig != "")

plot_df_sig <- mean_abund_sig |>
  pivot_longer(
    cols = c(corncob, woodchip),
    names_to = "SampleType",
    values_to = "rel_abund"
  ) |>
  mutate(
    rel_abund = ifelse(SampleType == "corncob", -rel_abund, rel_abund)
  )

length(unique(plot_df_sig$Phylum)) # 317 ASVs --> 133 Genus --> 20 Phyla

sig_dif_317 <- unique(plot_df_sig$OTU)

# NOTE: Reason why relative abundance counts are so low are because: The relative abundance number (like 6.1201e-03) means that across all samples in that group, the taxon has on average about 0.006 (or 0.6%) abundance - This reflects the natural variability in microbial communities: taxa can be very abundant in some samples and rare in others, causing the averaged relative abundance to be comparatively low.

#-----------------------------------------------------
# How many taxa are enriched in CC vs enriched in WC?
#-----------------------------------------------------

# separating ASVs greater in corn vs ASVs greater in wood chips
CC_greater <- mean_abund_sig |> filter(corncob > woodchip) |> distinct(OTU)
WC_greater <- mean_abund_sig |> filter(woodchip > corncob) |> distinct(OTU)

# count ASVs
nrow(CC_greater) # 71
nrow(WC_greater) # 246

# list of media-specific ETs
CC_enriched_taxa <- list(CC_greater$OTU)
WC_enriched_taxa <- list(WC_greater$OTU)

#--------------------------
# Log2Fold Change df prep
#--------------------------
pseudo <- 1e-6

log2fc_df <- mean_abund |>
  mutate(
    log2FC = log2((corncob + pseudo) / (woodchip + pseudo))
  )

sig_threshold <- 0.05

log2fc_df <- log2fc_df %>%
  mutate(
    result = case_when(
      p < sig_threshold & log2FC > 0 ~ "CC > WC",
      p < sig_threshold & log2FC < 0 ~ "CC < WC",
      TRUE ~ "ns"
    )
  )
```

*SUPP FINAL: Figure S4*
```{r, fig.width=6, fig.height=8}
sfig5 <- log2fc_df %>%
  ggplot(aes(x = log2FC, y = -log10(p), color = result)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_vline(xintercept = c(0), linetype = "dashed", color = "gray") +
  geom_hline(yintercept = -log10(sig_threshold), linetype = "dashed", color = "grey") +
  # geom_segment(aes(x = 0, y = -log10(0.05), xend = 0, yend = 13.21173), color = "red", linetype = "dashed") +
  # geom_segment(aes(x = 0, y = -log10(0.05), xend = 18.5, yend = -log10(0.05)), color = "red", linetype = "dashed") +
  scale_color_manual(
    values = c(
      "CC > WC" = "#ffa52d",        
      "CC < WC" = "darkseagreen",
      "not significant" = "grey70"
    ),
    name = "ASV enriched in",
    labels = c(
      "CC > WC" = "CC",        
      "CC < WC" = "WC"
    )
  ) +
  labs(
    x = "log2 Fold Change",
    y = "-log10(p-value)"
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 20),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 20),
    legend.title = element_text(size = 20)
  )

#ggsave("../raw_figs_301/dnbr_sfig5_volcano.png", plot = sfig5, width = 8, height = 6, units = "in", dpi = 300)
```

*Figure 3A*
```{r, fig.width=6, fig.height=15}
# se-ASVs only
log2fc_df <- log2fc_df |> filter(sig != "")

fig3a <- log2fc_df |>
  ggplot(aes(x = reorder(OTU, log2FC), y = log2FC, fill = log2FC > 0)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "#ffa52d", "FALSE" = "darkseagreen"),
    labels = c("TRUE" = "CC", "FALSE" = "WC"),
    name = "Enriched in"
  ) +
  labs(
    title = "16-h HRT",
    y = "log2 Fold Change", 
    x = "Each row = Unique se-ASV (317)") +
  theme_classic() +
  theme(axis.text.y = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 25),
        axis.text.x = element_text(size = 25),
        axis.title = element_text(size = 25),
        legend.title = element_text(size = 25),
        title = element_text(size = 25))
```

*ASV ORDER*
```{r}
asv_order_317 <- log2fc_df$OTU[order(log2fc_df$log2FC)]
```

*################################################################################################################*
*################################################################################################################*
#-----------------------------------
### 4.5) P/A heatmaps - (Fig3A)
# 317 media samples only, all HRTs
#-----------------------------------

*Filtering for 317 se-ASVs*
```{r}
se_asv_317 <- c("ASV1", "ASV10", "ASV101", "ASV102", "ASV103", "ASV104", "ASV105", "ASV1058", "ASV106", "ASV107", "ASV109", "ASV11", "ASV110", "ASV1107", "ASV1109", 
                "ASV111", "ASV113", "ASV114", "ASV1156", "ASV1158", "ASV117", "ASV118", "ASV119", "ASV12", "ASV120", "ASV123", "ASV127", "ASV129", "ASV13", "ASV131", 
                "ASV133", "ASV135", "ASV136", "ASV139", "ASV14", "ASV141", "ASV142", "ASV143", "ASV145", "ASV146", "ASV149", "ASV150", "ASV154", "ASV155", "ASV156", 
                "ASV158", "ASV159", "ASV16", "ASV162", "ASV164", "ASV167", "ASV168", "ASV169", "ASV17", "ASV170", "ASV171", "ASV172", "ASV175", "ASV179", "ASV18", 
                "ASV181", "ASV184", "ASV190", "ASV195", "ASV197", "ASV2", "ASV20", "ASV201", "ASV202", "ASV203", "ASV204", "ASV209", "ASV21", "ASV210", "ASV211", 
                "ASV212", "ASV213", "ASV214", "ASV217", "ASV221", "ASV222", "ASV227", "ASV228", "ASV23", "ASV231", "ASV232", "ASV234", "ASV237", "ASV238", "ASV24", 
                "ASV240", "ASV241", "ASV245", "ASV248", "ASV249", "ASV25", "ASV251", "ASV26", "ASV262", "ASV263", "ASV264", "ASV265", "ASV268", "ASV271", "ASV275", 
                "ASV277", "ASV280", "ASV285", "ASV289", "ASV291", "ASV292", "ASV293", "ASV295", "ASV296", "ASV298", "ASV299", "ASV30", "ASV300", "ASV303", "ASV305", 
                "ASV307", "ASV31", "ASV312", "ASV314", "ASV315", "ASV316", "ASV317", "ASV318", "ASV319", "ASV32", "ASV323", "ASV324", "ASV326", "ASV329", "ASV330", 
                "ASV331", "ASV333", "ASV334", "ASV335", "ASV337", "ASV338", "ASV34", "ASV340", "ASV341", "ASV344", "ASV349", "ASV351", "ASV354", "ASV359", "ASV36", 
                "ASV361", "ASV362", "ASV364", "ASV370", "ASV373", "ASV379", "ASV38", "ASV380", "ASV382", "ASV385", "ASV389", "ASV39", "ASV390", "ASV391", "ASV393", 
                "ASV395", "ASV399", "ASV4", "ASV40", "ASV400", "ASV405", "ASV407", "ASV41", "ASV415", "ASV416", "ASV419", "ASV421", "ASV422", "ASV423", "ASV425", 
                "ASV43", "ASV430", "ASV433", "ASV435", "ASV437", "ASV44", "ASV440", "ASV442", "ASV447", "ASV45", "ASV455", "ASV456", "ASV46", "ASV465", "ASV47", 
                "ASV470", "ASV471", "ASV478", "ASV48", "ASV484", "ASV486", "ASV490", "ASV491", "ASV499", "ASV5", "ASV50", "ASV501", "ASV502", "ASV503", "ASV504", 
                "ASV507", "ASV509", "ASV518", "ASV519", "ASV52", "ASV531", "ASV534", "ASV54", "ASV540", "ASV542", "ASV543", "ASV546", "ASV549", "ASV552", "ASV56", 
                "ASV564", "ASV565", "ASV568", "ASV570", "ASV579", "ASV58", "ASV581", "ASV582", "ASV584", "ASV587", "ASV588", "ASV589", "ASV59", "ASV594", "ASV596", 
                "ASV598", "ASV6", "ASV60", "ASV600", "ASV617", "ASV62", "ASV621", "ASV635", "ASV638", "ASV64", "ASV640", "ASV642", "ASV646", "ASV65", "ASV651", "ASV652", 
                "ASV670", "ASV672", "ASV673", "ASV675", "ASV679", "ASV68", "ASV681", "ASV682", "ASV686", "ASV690", "ASV693", "ASV7", "ASV700", "ASV703", "ASV705", "ASV71", 
                "ASV714", "ASV717", "ASV718", "ASV720", "ASV728", "ASV73", "ASV732", "ASV736", "ASV738", "ASV74", "ASV748", "ASV751", "ASV763", "ASV771", "ASV783", "ASV79", 
                "ASV798", "ASV8", "ASV806", "ASV809", "ASV810", "ASV811", "ASV82", "ASV825", "ASV826", "ASV835", "ASV838", "ASV842", "ASV843", "ASV85", "ASV86", "ASV87", 
                "ASV871", "ASV88", "ASV886", "ASV89", "ASV9", "ASV90", "ASV907", "ASV92", "ASV94", "ASV958", "ASV967", "ASV98", "ASV99" )

# prune taxa
ps_media_subset_317 <- prune_taxa(se_asv_317, ps_media)

# saving subset for PICRUSt2
#saveRDS(ps_media_subset_317, file = "../data/microbial/ps_media_317_114.rds") # 317 taxa, 111 samples

se_asv_317_list <- c("Xanthobacter", "Other", "Pseudorhodoplanes", "Azotobacter", 
  "Burkholderia-Caballeronia-Paraburkholderia", "Cellulomonas", "Azospira", 
  "Zoogloea", "Telmatospirillum", "Pseudomonas", "Plot4-2H12", "Afipia", 
  "Methanobacterium", "Cytophaga", "Methylophilus", "Taibaiella", 
  "Methanocella", "Mycobacterium", "Bradyrhizobium", "Nevskia", 
  "Pir2 lineage", "Ellin516", "Sphingomonas", "Geothrix", "Tahibacter", 
  "Desulfovibrio", "Bryobacter", "Mesorhizobium", "Methanomassiliicoccus", 
  "Rhizomicrobium", "Caulobacter", "Methylocystis", "Pleomorphomonas", 
  "Coxiella", "Sediminibacterium", "Bacillus", "Pajaroellobacter", 
  "Undibacterium", "Edaphobaculum", "Pseudolabrys", "Kaistia", 
  "Hyphomicrobium", "Rudaea", "Cupriavidus", "Devosia", "Methanosarcina", 
  "Sphingobium", "Chitinophaga", "Aetherobacter", "Altererythrobacter", 
  "Prosthecobacter", "Beijerinckia", "Leptolinea", "Terrimicrobium", 
  "SN8", "Heliimonas", "Aquisphaera", "Paludisphaera", "Fimbriiglobus", 
  "Microvirgula", "Opitutus", "Chthoniobacter", "Aquicella", "Puia", 
  "Novosphingobium", "Papillibacter", "Ruminiclostridium", "Dyella", 
  "LD29", "Anaerolineaceae UCG-001", "Bacteroides", "Pseudoxanthobacter", 
  "WCHB1-32", "Terrimonas", "Anaerospora", "Fibrisoma", "Thiobacillus", 
  "Curvibacter", "Pedomicrobium", "Solimonas", "Magnetospirillum", 
  "Aminobacter", "Desulfosporosinus", "Streptomyces", "Parachlamydia", 
  "Clostridium sensu stricto 1", "SM1A02", "NK4A214 group", "Ellin6055", 
  "Gluconacetobacter", "Incertae Sedis", "Sphingoaurantiacus", 
  "Singulisphaera", "Haliangium", "Christensenellaceae R-7 group", 
  "Gemmata", "SH3-11", "Rhodopila", "Anaeromyxobacter", 
  "Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium", "Rhodoplanes", 
  "Clostridium sensu stricto 12", "Labrys", "Azospirillum", "Inquilinus", 
  "Steroidobacter", "Gordonia", "Thermomonas", "Nordella", "Rhodovastum", 
  "Psychrobacter", "Mucilaginibacter", "Galbitalea", "Methanospirillum", 
  "Acidibacter", "Ferruginibacter", "Thermincola", "Limnohabitans", 
  "Sporobacter", "Williamsia", "Lacunisphaera", 
  "Defluviitaleaceae UCG-011", "Reyranella", "Bdellovibrio", 
  "Anaerolinea", "Spirosoma", "Gemmatimonas", 
  "Methylobacterium-Methylorubrum", "Sporomusa", "Brevundimonas", 
  "Tundrisphaera", "Pseudobacteroides", "Spirochaeta 2")
```

*Figure 3A - 4-h HRT p/a heatmap (& prep)*
```{r, fig.width=10, fig.height=15}
# subsetting 
ps_4HRT <- subset_samples(ps16S_final, HRT == "4")
ps_4HRT_media <- subset_samples(ps_4HRT, SampleSource == "media")
ps_4_317 <- prune_taxa(se_asv_317, ps_4HRT_media)
ps_4_317_df <- data.frame(sample_data(ps_4_317), stringsAsFactors = FALSE)

# correct order of sample IDs - #unique(ps_4_317_df$SampleID) 
HRT4_sampID_order <- c(
"4CC_1_0", "4CC_1_7", "4CC_1_11", "4CC_1_14", "4CC_1_18", 
"4CC_2_0", "4CC_2_4", "4CC_2_7", "4CC_2_11", "4CC_2_14", "4CC_2_18", "4CC_2_21", 
"4CC_3_0", "4CC_3_4", "4CC_3_7", "4CC_3_11", "4CC_3_14", "4CC_3_18", 
"4WC_4_4", "4WC_4_8", "4WC_4_11", "4WC_4_14", "4WC_4_21",  
"4WC_5_0", "4WC_5_4", "4WC_5_11", "4WC_5_14", "4WC_5_21",  
"4WC_6_0", "4WC_6_4", "4WC_6_11", "4WC_6_14", "4WC_6_18", "4WC_6_21")

ps_4_317_df$SampleID <- factor(ps_4_317_df$SampleID, levels = HRT4_sampID_order)
sample_data(ps_4_317) <- ps_4_317_df

# relative abundance calculation
ps_4_317_rel <- transform_sample_counts(ps_4_317, function(x) x / sum(x))

# melt pso
ps_4_317_rel_abund_melt <- psmelt(ps_4_317_rel) %>%
  filter(!is.na(OTU), OTU != "NA")

# presence/absence
ps_4_317_pa <- ps_4_317_rel_abund_melt |>
  mutate(Present = ifelse(Abundance > 0, 1, 0))

# getting grouping from log2fc_df 
log2fc_df_labels <- log2fc_df |>
  dplyr::select(OTU, result)

ps_4_317_pa <- left_join(ps_4_317_pa, log2fc_df_labels, by = "OTU")

ps_4_317_side_annot_df <- ps_4_317_pa %>%
  distinct(OTU, result) %>%
  mutate(x = "4-h HRT ID")

ps_4_317_side_annot_df$SampleType <- "corncob" # so annot is only on corncob facted plot, not WC

ps_4_317_pa$OTU <- factor(ps_4_317_pa$OTU, levels = asv_order_317)

# custom strip label
HRT4_hm_strips <- c(
  "corncob" = "CC", 
  "woodchip" = "WC")

# plot ------------------------------------------------------------------------------------------------
pa_4_plot <- ggplot() +
  geom_tile(data = ps_4_317_side_annot_df,
            aes(x = x, y = OTU, fill = result),
            width = 0.7) +
  scale_fill_manual(
    name = "Enriched in",
    values = c("CC > WC" = "#ffa52d", "CC < WC" = "darkseagreen"),
    guide = "none"
  ) +
  labs(
    title = "4-h HRT",
    x = "Each column = One sample",
    y = "Each row = Unique se-ASV (317)"
  ) +
  new_scale_fill() +
  geom_tile(data = ps_4_317_pa %>% filter(SampleType == "corncob"),
            aes(x = SampleID, y = OTU, fill = factor(Present)),
            color = NA) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"),
                    name = "Presence", guide = "none") +
  new_scale_fill() +
  geom_tile(data = ps_4_317_pa %>% filter(SampleType == "woodchip"),
            aes(x = SampleID, y = OTU, fill = factor(Present)),
            color = NA) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"),
                    name = "Presence", guide = "none") +
  facet_wrap2(~ SampleType,
              scales = "free_x",
              labeller = as_labeller(HRT4_hm_strips)) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    axis.text.x  = element_blank(),
    axis.text.y  = element_blank(),
    axis.title.x = element_text(size = 25),
    axis.title.y = element_text(size = 25),
    strip.text = element_text(size = 25),
    title = element_text(size = 25)
  )
```

*Figure 3A - 8-h HRT p/a heatmap (& prep)*
```{r, fig.width=10, fig.height=15}
# subsetting 
ps_8HRT <- subset_samples(ps16S_final, HRT == "8")
ps_8HRT_media <- subset_samples(ps_8HRT, SampleSource == "media")
ps_8_317 <- prune_taxa(se_asv_317, ps_8HRT_media)
ps_8_317_df <- data.frame(sample_data(ps_8_317), stringsAsFactors = FALSE)

# correct order of sample IDs - #unique(ps_8_317_df$SampleID) 

HRT8_sampID_order <- c(
 "CC_1_6", "CC_1_10", "CC_1_15", "CC_1_20",
 "CC_2_6", "CC_2_15", "CC_2_20",
 "CC_3_6", "CC_3_10", "CC_3_15", "CC_3_20",
 "WC_4_6", "WC_4_10", "WC_4_15", "WC_4_20",
 "WC_5_6", "WC_5_10", "WC_5_15", "WC_5_20",
 "WC_6_6", "WC_6_10", "WC_6_15", "WC_6_20"
)

ps_8_317_df$SampleID <- factor(ps_8_317_df$SampleID, levels = HRT8_sampID_order)
sample_data(ps_8_317) <- ps_8_317_df

# relative abundance calculation
ps_8_317_rel <- transform_sample_counts(ps_8_317, function(x) x / sum(x))

# melt pso
ps_8_317_rel_abund_melt <- psmelt(ps_8_317_rel) %>%
  filter(!is.na(OTU), OTU != "NA")

# presence/absence
ps_8_317_pa <- ps_8_317_rel_abund_melt |>
  mutate(Present = ifelse(Abundance > 0, 1, 0))

# getting grouping from log2fc_df 
log2fc_df_labels <- log2fc_df |>
  dplyr::select(OTU, result)

ps_8_317_pa <- left_join(ps_8_317_pa, log2fc_df_labels, by = "OTU")

ps_8_317_side_annot_df <- ps_8_317_pa %>%
  distinct(OTU, result) %>%
  mutate(x = "8-h HRT ID")

ps_8_317_side_annot_df$SampleType <- "corncob" # so annot is only on corncob facted plot, not WC

ps_8_317_pa$OTU <- factor(ps_8_317_pa$OTU, levels = asv_order_317)

# custom strip label
HRT8_hm_strips <- c(
  "corncob" = "CC", 
  "woodchip" = "WC")

# plot ------------------------------------------------------------------------------------------------
pa_8_plot <- ggplot() +
  geom_tile(data = ps_8_317_side_annot_df,
            aes(x = x, y = OTU, fill = result),
            width = 0.5) +
  scale_fill_manual(
    name = "Enriched in",
    values = c("CC > WC" = "#ffa52d", "CC < WC" = "darkseagreen"),
    labels = c("CC > WC" = "CC", "CC < WC" = "WC")
  ) +
  labs(
    title = "8-h HRT",
    x = "Each column = One sample",
    y = "Each row = Unique se-ASV (317)"
  ) +
  new_scale_fill() +
  geom_tile(data = ps_8_317_pa %>% filter(SampleType == "corncob"),
            aes(x = SampleID, y = OTU, fill = factor(Present)),
            color = NA) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"),
                    name = "Presence", guide = "none") +
  new_scale_fill() +
  geom_tile(data = ps_8_317_pa %>% filter(SampleType == "woodchip"),
            aes(x = SampleID, y = OTU, fill = factor(Present)),
            color = NA) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"),
                    name = "Presence", guide = "none") +
  facet_wrap2(~ SampleType,
              scales = "free_x",
              labeller = as_labeller(HRT8_hm_strips)) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    axis.text.x  = element_blank(),
    axis.text.y  = element_blank(),
    axis.title.x = element_text(size = 25),
    axis.title.y = element_text(size = 25),
    strip.text = element_text(size = 25),
    legend.position = "bottom",
    legend.text = element_text(size = 25),
    title = element_text(size = 25)
  )
```

*Figure 3A - 16-h HRT p/a heatmap (& prep)*
```{r, fig.width=10, fig.height=15}
# subsetting
ps_16HRT <- subset_samples(ps16S_final, HRT == "16")
ps_16HRT_media <- subset_samples(ps_16HRT, SampleSource == "media")
ps_16_317 <- prune_taxa(se_asv_317, ps_16HRT_media)
ps_16_317_df <- data.frame(sample_data(ps_16_317), stringsAsFactors = FALSE)

# correct order of sample IDs - #unique(ps_16_317_df$SampleID) 

HRT16_sampID_order <- c(
"CC_1_21", "CC_1_28", "CC_1_29", "CC_1_30", "CC_1_31", "CC_1_32", "CC_1_33", "CC_1_34", "CC_1_35", "CC_1_36", "CC_1_37", "CC_1_38", "CC_2_21",
"CC_2_28", "CC_2_29", "CC_2_30", "CC_2_31", "CC_2_32", "CC_2_33", "CC_2_34", "CC_2_35", "CC_2_36", "CC_2_37", "CC_2_38", "CC_3_21", "CC_3_28",
"CC_3_29", "CC_3_30", "CC_3_31", "CC_3_32", "CC_3_33", "CC_3_34", "CC_3_35", "CC_3_36", "CC_3_37", "CC_3_38", "WC_4_21", "WC_4_24", "WC_4_27",
"WC_4_28", "WC_4_29", "WC_4_30", "WC_4_31", "WC_5_21", "WC_5_24", "WC_5_27", "WC_5_28", "WC_5_29", "WC_5_30", "WC_5_31", "WC_6_21", "WC_6_24",
"WC_6_27", "WC_6_28", "WC_6_29", "WC_6_30", "WC_6_31"
)

ps_16_317_df$SampleID <- factor(ps_16_317_df$SampleID, levels = HRT16_sampID_order)
sample_data(ps_16_317) <- ps_16_317_df

# relative abundance calculation 
ps_16_317_rel <- transform_sample_counts(ps_16_317, function(x) x / sum(x))

# melt pso
ps_16_317_rel_abund_melt <- psmelt(ps_16_317_rel) %>%
  filter(!is.na(OTU), OTU != "NA")

# presence/absence
ps_16_317_pa <- ps_16_317_rel_abund_melt |>
  mutate(Present = ifelse(Abundance > 0, 1, 0))

# getting grouping from log2fc_df 
log2fc_df_labels <- log2fc_df |>
  dplyr::select(OTU, result)

ps_16_317_pa <- left_join(ps_16_317_pa, log2fc_df_labels, by = "OTU")

ps_16_317_side_annot_df <- ps_16_317_pa %>%
  distinct(OTU, result) %>%
  mutate(x = "16-h HRT ID")

ps_16_317_side_annot_df$SampleType <- "corncob" # so annot is only on corncob facted plot, not WC

ps_16_317_pa$OTU <- factor(ps_16_317_pa$OTU, levels = asv_order_317)

# custom strip label
HRT16_hm_strips <- c(
  "corncob" = "CC", 
  "woodchip" = "WC")

# plot ------------------------------------------------------------------------------------------------
pa_16_plot <- ggplot() +
  geom_tile(data = ps_16_317_side_annot_df,
            aes(x = x, y = OTU, fill = result),
            width = 1) +
  scale_fill_manual(
    name = "Enriched in",
    values = c("CC > WC" = "#ffa52d", "CC < WC" = "darkseagreen"),
    guide = "none"
  ) +
  labs(
    title = "16-h HRT",
    x = "Each column = One sample",
    y = "Each row = Unique se-ASV (317)"
  ) +
  new_scale_fill() +
  geom_tile(data = ps_16_317_pa %>% filter(SampleType == "corncob"),
            aes(x = SampleID, y = OTU, fill = factor(Present)),
            color = NA) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"),
                    name = "Presence") +
  new_scale_fill() +
  geom_tile(data = ps_16_317_pa %>% filter(SampleType == "woodchip"),
            aes(x = SampleID, y = OTU, fill = factor(Present)),
            color = NA) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"),
                    name = "Presence") +
  facet_wrap2(~ SampleType,
              scales = "free_x",
              labeller = as_labeller(HRT16_hm_strips)) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    axis.text.x  = element_blank(),
    axis.text.y  = element_blank(),
    axis.title.x = element_text(size = 25),
    axis.title.y = element_text(size = 25),
    strip.text = element_text(size = 25),
    legend.position = "bottom",
    legend.text = element_text(size = 25),
    title = element_text(size = 25)
  )
```

*FINAL MAIN: Figure 3A*
```{r, fig.width=30, fig.height=15}
combined_heatmap <- plot_spacer() + fig3a + pa_16_plot + pa_8_plot + pa_4_plot +
  plot_layout(ncol = 5, widths = c(0.05, 0.75, 1, 1, 1))

combined_heatmap

#ggsave("../raw_figs_301/dnbr_fig3a.png", plot = combined_heatmap, width = 20, height = 12, units = "in", dpi = 300)
```

*################################################################################################################*
*################################################################################################################*
#--------------------------------------------------------
### 4.75) Summed relative abundance of se-ASVs - (FigS5)
# Media samples only, all HRTs
#--------------------------------------------------------

*se-ASV lists*
```{r}
CC_71_se_asvs <- c("ASV1", "ASV10", "ASV101", "ASV105", "ASV106", "ASV107", "ASV109", "ASV11", "ASV12", "ASV13", "ASV146", "ASV149", "ASV150", "ASV154", "ASV156", 
                   "ASV158", "ASV16", "ASV171", "ASV172", "ASV18", "ASV184", "ASV2", "ASV20", "ASV204", "ASV21", "ASV213", "ASV217", "ASV221", "ASV24", "ASV240", 
                   "ASV241", "ASV245", "ASV249", "ASV25", "ASV262", "ASV275", "ASV293", "ASV307", "ASV31", "ASV312", "ASV32", "ASV326", "ASV330", "ASV337", "ASV34", 
                   "ASV341", "ASV36", "ASV395", "ASV4", "ASV40", "ASV423", "ASV44", "ASV46", "ASV484", "ASV490", "ASV54", "ASV570", "ASV58", "ASV59", "ASV6", "ASV690", 
                   "ASV7", "ASV71", "ASV73", "ASV74", "ASV85", "ASV88", "ASV89", "ASV9", "ASV90", "ASV99" ) 

WC_246_se_asvs <- c("ASV102", "ASV103", "ASV104", "ASV1058", "ASV110", "ASV1107", "ASV1109", "ASV111", "ASV113", "ASV114", "ASV1156", "ASV1158", "ASV117", "ASV118", 
                    "ASV119", "ASV120", "ASV123", "ASV127", "ASV129", "ASV131", "ASV133", "ASV135", "ASV136", "ASV139", "ASV14", "ASV141", "ASV142", "ASV143", "ASV145", 
                    "ASV155", "ASV159", "ASV162", "ASV164", "ASV167", "ASV168", "ASV169", "ASV17", "ASV170", "ASV175", "ASV179", "ASV181", "ASV190", "ASV195", "ASV197", 
                    "ASV201", "ASV202", "ASV203", "ASV209", "ASV210", "ASV211", "ASV212", "ASV214", "ASV222", "ASV227", "ASV228", "ASV23", "ASV231", "ASV232", "ASV234", 
                    "ASV237", "ASV238", "ASV248", "ASV251", "ASV26", "ASV263", "ASV264", "ASV265", "ASV268", "ASV271", "ASV277", "ASV280", "ASV285", "ASV289", "ASV291", 
                    "ASV292", "ASV295", "ASV296", "ASV298", "ASV299", "ASV30", "ASV300", "ASV303", "ASV305", "ASV314", "ASV315", "ASV316", "ASV317", "ASV318", "ASV319", 
                    "ASV323", "ASV324", "ASV329", "ASV331", "ASV333", "ASV334", "ASV335", "ASV338", "ASV340", "ASV344", "ASV349", "ASV351", "ASV354", "ASV359", "ASV361", 
                    "ASV362", "ASV364", "ASV370", "ASV373", "ASV379", "ASV38", "ASV380", "ASV382", "ASV385", "ASV389", "ASV39", "ASV390", "ASV391", "ASV393", "ASV399", 
                    "ASV400", "ASV405", "ASV407", "ASV41", "ASV415", "ASV416", "ASV419", "ASV421", "ASV422", "ASV425", "ASV43", "ASV430", "ASV433", "ASV435", "ASV437", 
                    "ASV440", "ASV442", "ASV447", "ASV45", "ASV455", "ASV456", "ASV465", "ASV47", "ASV470", "ASV471", "ASV478", "ASV48", "ASV486", "ASV491", "ASV499", 
                    "ASV5", "ASV50", "ASV501", "ASV502", "ASV503", "ASV504", "ASV507", "ASV509", "ASV518", "ASV519", "ASV52", "ASV531", "ASV534", "ASV540", "ASV542", 
                    "ASV543", "ASV546", "ASV549", "ASV552", "ASV56", "ASV564", "ASV565", "ASV568", "ASV579", "ASV581", "ASV582", "ASV584", "ASV587", "ASV588", "ASV589", 
                    "ASV594", "ASV596", "ASV598", "ASV60", "ASV600", "ASV617", "ASV62", "ASV621", "ASV635", "ASV638", "ASV64", "ASV640", "ASV642", "ASV646", "ASV65", 
                    "ASV651", "ASV652", "ASV670", "ASV672", "ASV673", "ASV675", "ASV679", "ASV68", "ASV681", "ASV682", "ASV686", "ASV693", "ASV700", "ASV703", "ASV705", 
                    "ASV714", "ASV717", "ASV718", "ASV720", "ASV728", "ASV732", "ASV736", "ASV738", "ASV748", "ASV751", "ASV763", "ASV771", "ASV783", "ASV79", "ASV798", 
                    "ASV8", "ASV806", "ASV809", "ASV810", "ASV811", "ASV82", "ASV825", "ASV826", "ASV835", "ASV838", "ASV842", "ASV843", "ASV86", "ASV87", "ASV871", 
                    "ASV886", "ASV907", "ASV92", "ASV94", "ASV958", "ASV967", "ASV98") 
```

*Corncob: proportion of se-ASVs to total ASV community*
```{r, fig.width=5, fig.height=7}
# pso subset
ps_media_cc <- subset_samples(ps_media, SampleType == "corncob")

# pruning 0 abundance ASVs across all samples
ps_media_cc_otu_mat <- as(otu_table(ps_media_cc), "matrix")
non0_asvs <- colSums(ps_media_cc_otu_mat) > 0
ps_media_pruned <- prune_taxa(non0_asvs, ps_media_cc)

# calculating relative abundance within Sample Source, HRT, and Bioreactor
ps_media_cc_rel_df <- ps_media_pruned |>
  psmelt() |>
  as_tibble() |>
  group_by(Sample, HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# se-ASVs vs. non-se-ASVs
CC_71_se_asvs
cc_asvs <- unique(ps_media_cc_rel_df$OTU) # 639 (all HRTs)
CC_non_se_asvs <- setdiff(cc_asvs, CC_71_se_asvs) # 568

# adding classification column for se_asv vs. non_se_asvs
ps_media_cc_rel_df <- ps_media_cc_rel_df |>
  mutate(ASV_Group = if_else(OTU %in% CC_71_se_asvs, "CC_se_ASV", "CC_non_se_ASV"))

# For each sample, sum relative abundance of CC_se_ASVs per HRT (and likewise for non_se_ASVs if desired)
cc_se_sample_relab <- ps_media_cc_rel_df %>%
  filter(ASV_Group == "CC_se_ASV") %>%
  group_by(Sample, HRT) %>%
  summarise(total_cc_se_asv_rel_abund = sum(relative_abund, na.rm = TRUE), .groups = "drop")

cc_se_sample_relab$SampleType <- sapply(cc_se_sample_relab$Sample, extract_sample_type)

# kruskal
cc_se_sample_relab |>
  kruskal_test(total_cc_se_asv_rel_abund ~ HRT) # p = 1.02e-10

# Dunn
cc_se_sample_dunn <- cc_se_sample_relab |>
  dunn_test(total_cc_se_asv_rel_abund ~ HRT, p.adjust.method = "bonferroni")

# plot
cc_box <- ggplot(cc_se_sample_relab, aes(x = HRT, y = total_cc_se_asv_rel_abund, fill = SampleType)) +
  geom_boxplot() +
#  geom_jitter(width = 0.2, alpha = 0.5) +
  scale_fill_manual(
    labels = c("corncob" = "CC"),
    values = c("corncob" = "#ffa52d"),
    name = "Media Type") +
  theme_classic() +
  labs(x = "HRT", y = "Summed Relative Abundance of CC_se_ASVs") +
  geom_signif(
    comparisons = list(c("4", "8"), c("8", "16"), c("4", "16")),
    annotations = c("ns", "**", "****"),
    y_position = c(1.00, 1.05, 1.1),   # adjust to be above your highest value
    tip_length = 0.03,
    textsize = 6
  ) +
  theme(axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        legend.position = "bottom",
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15))
```

*Woodchip: proportion of se-ASVs to total ASV community*
```{r, fig.width=5, fig.height=7}
# pso subset
ps_media_WC <- subset_samples(ps_media, SampleType == "woodchip")

# pruning 0 abundance ASVs across all samples
ps_media_WC_otu_mat <- as(otu_table(ps_media_WC), "matrix")
non0_asvs <- colSums(ps_media_WC_otu_mat) > 0
ps_media_pruned <- prune_taxa(non0_asvs, ps_media_WC)

# calculating relative abundance within Sample Source, HRT, and Bioreactor
ps_media_WC_rel_df <- ps_media_pruned |>
  psmelt() |>
  as_tibble() |>
  group_by(Sample, HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# se-ASVs vs. non-se-ASVs
WC_246_se_asvs
wc_asvs <- unique(ps_media_WC_rel_df$OTU) # 1282 (all HRTs)
wc_non_se_asvs <- setdiff(wc_asvs, WC_246_se_asvs) # 1036

# adding classification column for se_asv vs. non_se_asvs
ps_media_WC_rel_df <- ps_media_WC_rel_df |>
  mutate(ASV_Group = if_else(OTU %in% WC_246_se_asvs, "WC_se_ASV", "WC_non_se_ASV"))

# For each sample, sum relative abundance of WC_se_ASVs per HRT (and likewise for non_se_ASVs if desired)
WC_se_sample_relab <- ps_media_WC_rel_df %>%
  filter(ASV_Group == "WC_se_ASV") %>%
  group_by(Sample, HRT) %>%
  summarise(total_WC_se_asv_rel_abund = sum(relative_abund, na.rm = TRUE), .groups = "drop")

WC_se_sample_relab$SampleType <- sapply(WC_se_sample_relab$Sample, extract_sample_type)

# kruskal
WC_se_sample_relab |>
  kruskal_test(total_WC_se_asv_rel_abund ~ HRT) # p = 2.39e-08

# Dunn
WC_se_sample_dunn <- WC_se_sample_relab |>
  dunn_test(total_WC_se_asv_rel_abund ~ HRT, p.adjust.method = "bonferroni")

# plot
wc_box <- ggplot(WC_se_sample_relab, aes(x = HRT, y = total_WC_se_asv_rel_abund, fill = SampleType)) +
  geom_boxplot() +
#  geom_jitter(width = 0.2, alpha = 0.5) +
  scale_fill_manual(
    labels = c("woodchip" = "WC"),
    values = c("woodchip" = "darkseagreen"),
    name = "Media Type") +
  theme_classic() +
  labs(x = "HRT", y = "Summed Relative Abundance of WC_se_ASVs") +
  geom_signif(
    comparisons = list(c("4", "8"), c("8", "16"), c("4", "16")),
    annotations = c("**", "ns", "****"),
    y_position = c(1.00, 1.05, 1.1),   # adjust to be above your highest value
    tip_length = 0.03,
    textsize = 6
  ) + 
  theme(axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        legend.position = "bottom",
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))
```

*FINAL SUPP: se-ASV rel abund box figure*
```{r, fig.width=10, fig.height=7}
box_p <- cc_box + plot_spacer() + wc_box + 
  plot_layout(widths = c(1, 0.1, 1), ncol = 3) 

# save plot
#ggsave("../raw_figs_301/dnbr_sfig5_se_asv_rel_abund.png", plot = box_p, width = 8, height = 6, units = "in", dpi = 300)
```

*################################################################################################################*
*################################################################################################################*
#----------------------------------------
### 5) Correlation Analysis - (Fig3B-C)
# Media samples only, all HRTs
#----------------------------------------

*NRE data in correlation format*
```{r}
nre_df <- read_csv("../data/performance/full_NRE_237_data_DBRLC_2026.csv")

nre_df <- nre_df |> dplyr::rename(
  SampleType = sample_type,
  Bioreactor = bioreactor_ID,
  Week = influent_id)

nre_df$SampleID <- ifelse(
  nre_df$HRT == 4,
  paste0("4", paste(nre_df$SampleType, nre_df$Bioreactor, nre_df$Week, sep = "_")),
  paste(nre_df$SampleType, nre_df$Bioreactor, nre_df$Week, sep = "_")
)

nre_df <- nre_df |>
  dplyr::select(SampleID, SampleType, Bioreactor, HRT, Week, removal_efficiency, TOC_conc)

# converting cols to factors
columns_to_factor <- c("SampleType", "HRT", "Bioreactor")

nre_df <- nre_df |>
  mutate_at(vars(all_of(columns_to_factor)), ~ as.factor(as.character(.)))

# ordering HRT
HRT_order <- c("4", "8", "16")

nre_df <- nre_df |>
  mutate(HRT = factor(HRT, levels = HRT_order))
```

*Prune taxa prep for CC & WC*
```{r}
# CC PSO prune taxa prep
#------------------------
ps_media_CC <- subset_samples(ps_media, SampleType == "corncob")

# transform phyloseq counts to relative abundances per sample <- ALL TAXA
ps_media_CC_rel <- transform_sample_counts(ps_media_CC, function(x) x / sum(x))

# extract OTU table as data.frame (samples as rows) 
ps_media_CC_otu <- as.data.frame(otu_table(ps_media_CC_rel))

# WC PSO prune taxa prep
#------------------------
ps_media_WC <- subset_samples(ps_media, SampleType == "woodchip")

# transform phyloseq counts to relative abundances per sample <- ALL TAXA
ps_media_WC_rel <- transform_sample_counts(ps_media_WC, function(x) x / sum(x))

# extract OTU table as data.frame (samples as rows)
ps_media_WC_otu <- as.data.frame(otu_table(ps_media_WC_rel))
```

*Figure 3B - all CC se-ASVs*
```{r, fig.width=6, fig.height=8}
ps_media_CC_71 <- ps_media_CC_otu |>
  dplyr::select(all_of(CC_71_se_asvs))

# sum relative abundances per sample across selected taxa 
ps_media_CC_71$SumTaxa <- rowSums(ps_media_CC_71, na.rm = TRUE)
ps_media_CC_71 <- ps_media_CC_71 |> dplyr::select(SumTaxa, everything())

# add SampleID column for merging
ps_media_CC_71$SampleID <- rownames(ps_media_CC_71)

# join with nitrate removal efficiency data frame (nre_df must have SampleID and removal_efficiency)
nre_CC_pso_df <- ps_media_CC_71 |>
  dplyr::select(SampleID, SumTaxa) |>
  inner_join(nre_df, by = "SampleID")

# plot - 71 CC se-ASVs & NRE -------------------------------------------------------------------------------------------------------------------
CC_71_corr_p <- nre_CC_pso_df |>
  ggplot(aes(x = removal_efficiency, y = SumTaxa, shape = HRT, colour = SampleType)) +
  geom_point(size = 3) +
  stat_cor(method = "spearman", 
           aes(label = paste0("rho = ", ..r.., ", p = ", formatC(..p.., format = "e", digits = 2)), group = 1),
           output.type = "text",
           label.x.npc = 0, label.y.npc = 0.95,
           size = 5, 
           color = "black") +
  labs(
    x = "Nitrate Removal Efficiency (%)",
    y = "Relative Abundance of CC se-ASVs (71)",
    color = "Sample Type",
#    title = "CC se-ASVs only"
  ) +
  theme_classic() +
  scale_color_manual(
    values = c("CC" = "black"), name = "Media Type"
  ) + 
  scale_shape_manual(
    values = c(
      "4" = 18, 
      "8" = 17,
      "16" = 16
    )
  ) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 20), 
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title.y = element_text(margin = margin(r = 10)))


# checking small p-value <- p = 1.402e-13, rho = 0.7873758 
cor.test(
  nre_CC_pso_df$removal_efficiency,
  nre_CC_pso_df$SumTaxa,
  method = "spearman",
  exact  = FALSE   
)
```

*Figure 3C - all WC se-ASVs*
```{r, fig.width=6, fig.height=8}
ps_media_WC_246 <- ps_media_WC_otu |>
  dplyr::select(all_of(WC_246_se_asvs))

# sum relative abundances per sample across selected taxa 
ps_media_WC_246$SumTaxa <- rowSums(ps_media_WC_246, na.rm = TRUE)
ps_media_WC_246 <- ps_media_WC_246 |> dplyr::select(SumTaxa, everything())

# add SampleID column for merging
ps_media_WC_246$SampleID <- rownames(ps_media_WC_246)

# join with nitrate removal efficiency data frame (nre_df must have SampleID and removal_efficiency)
nre_WC_pso_df <- ps_media_WC_246 |>
  dplyr::select(SampleID, SumTaxa) |>
  inner_join(nre_df, by = "SampleID")

# plot - 246 WC se-ASVs & NRE -------------------------------------------------------------------------------------------------------------------
WC_246_corr_p <- nre_WC_pso_df |>
  ggplot(aes(x = removal_efficiency, y = SumTaxa, shape = HRT, colour = SampleType)) +
  geom_point(size = 3) +
  stat_cor(method = "spearman", 
           aes(label = paste0("rho = ", ..r.., ", p = ", formatC(..p.., format = "e", digits = 2)), group = 1),
           output.type = "text",
           label.x.npc = 0.5, label.y.npc = 0.4,
           size = 5, 
           color = "black") +
  labs(
    x = "Nitrate Removal Efficiency (%)",
    y = "Relative Abundance of WC se-ASVs (246)",
    color = "Sample Type",
#    title = "WC se-ASVs only"
  ) +
  theme_classic() +
  scale_color_manual(
    values = c("WC" = "black"),
    name = "Media Type"
  ) + 
  scale_shape_manual(
    values = c(
      "4" = 18, 
      "8" = 17,
      "16" = 16
    )
  ) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 20), 
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title.y = element_text(margin = margin(r = 10)))


# checking small p-value <- p = 0.02242, rho = 0.3324468 
cor.test(
  nre_WC_pso_df$removal_efficiency,
  nre_WC_pso_df$SumTaxa,
  method = "spearman",
  exact  = FALSE  
)
```

*FINAL MAIN: Figure 3 B-C*
```{r, fig.width=14, fig.height=8}
se_ASVs_71_246_fig <- plot_spacer() + CC_71_corr_p + plot_spacer() + WC_246_corr_p +
  plot_layout(widths = c(0.05, 1, 0.1, 1), ncol = 4) 

# save plot
#ggsave("../raw_figs_301/dnbr_fig3b_c.png", plot = se_ASVs_71_246_fig, width = 14, height = 8, units = "in", dpi = 300)
```

*################################################################################################################*
*################################################################################################################*
#-------------------------------------------
# 6) PICRUSt2 - (Fig4, FigS5)
# 317 se-asvs, media samples only, 16-h HRT
#-------------------------------------------

*PICRUSt2 output files read in*
```{r}
ps_media_317_df <- sample_data(ps_media_subset_317) |>
  dplyr::as_tibble() |>
  as.data.frame()

rownames(ps_media_317_df) = ps_media_317_df$SampleID
sample_data(ps_media_subset_317) = data.frame(ps_media_317_df)

# ensure sample df is read in from above code
ps_media_317_df 

# read in files - with descriptions
KO_raw <- read_tsv("../data/picrust2/KO_317.tsv")

# files w/o descriptions <- since functions need numeric only data
KO_num <- read_tsv("../data/picrust2/KO_num_317.tsv")
```

*PICRUSt2 downstream prep*
```{r}
# converting raw abundance files 
KO_long <- function_to_long(KO_num)
KO_rel <- relative_abundance(KO_long)

# sample mapping <- all sample names should be identical
sample_info <- tibble(
  SampleID = unique(KO_rel$SampleID),
  SampleType = if_else(str_detect(SampleID, "CC"), "CC", 
           if_else(str_detect(SampleID, "WC"), "WC", NA_character_))
)

# adding rep & HRT
sample_info$Bioreactor <- sapply(sample_info$SampleID, extract_bioreactor_id)
sample_info$HRT <- sapply(sample_info$SampleID, extract_HRT)
sample_info$Time <- sapply(sample_info$SampleID, extract_week)

# converting cols to factors
columns_to_factor <- c("SampleType", "HRT", "Bioreactor", "Time")

sample_info <- sample_info |>
  mutate_at(vars(all_of(columns_to_factor)), ~ as.factor(as.character(.)))

# ordering HRT
HRT_order <- c("4", "8", "16")

sample_info <- sample_info |>
  mutate(HRT = factor(HRT, levels = HRT_order))

# retreiving descriptions from PICRUSt2 raw outputs
KO_descrip <- KO_raw |> dplyr::select(`function`, description)

# output files with sample information
KO_rel <- left_join(KO_rel, sample_info, by = "SampleID")

# output files with functional descriptions
KO_rel <- left_join(KO_rel, KO_descrip, by = "function")

#----------------
# KO Figure Prep
#----------------
# specific denitrification-associated K-numbers (source: https://www.kegg.jp/module/M00529)
KO_IDs <- c("ko:K00370", "ko:00371", "ko:K00374", "ko:K02567", "ko:K02568", "ko:K00368", "ko:K15864", "ko:K04561", "ko:K02305", "ko:K00376") 

DN_KO_rel <- KO_rel |> filter(`function` %in% KO_IDs)
DN_KO_rel <- DN_KO_rel |>
  mutate(`function` = factor(`function`, levels = KO_IDs))

# separating rel df by HRT
DN_KO_rel_4 <- DN_KO_rel |> filter(HRT == "4")
DN_KO_rel_8 <- DN_KO_rel |> filter(HRT == "8")
DN_KO_rel_16 <- DN_KO_rel |> filter(HRT == "16")

# custom stats annotations
plot_comps <- list(c("CC", "WC"))

# custom strips
KO_strip_text_labs <- c(
           "ko:K00370" = "K00370 (narG)", 
           "ko:K00374" = "K00374 (narI)", 
           "ko:K02567" = "K02567 (napA)", 
           "ko:K02568" = "K02568 (napB)", 
           "ko:K00368" = "K00368 (nirK)", 
           "ko:K15864" = "K15864 (nirS)", 
           "ko:K04561" = "K04561 (norB)", 
           "ko:K02305" = "K02305 (norC)", 
           "ko:K00376" = "K00376 (nosZ)"
           )

KO_strip_text_cols <- c(
  "K00370 (narG)" = "grey60", 
  "K00374 (narI)" = "grey60", 
  "K02567 (napA)" = "grey60", 
  "K02568 (napB)" = "grey60", 
  "K00368 (nirK)" = "grey80", 
  "K15864 (nirS)" = "grey80", 
  "K04561 (norB)" = "#aca193", 
  "K02305 (norC)" = "#aca193", 
  "K00376 (nosZ)" = "grey90"
)

strip <- strip_themed(
  text_x = elem_list_text(
    colour = "black"
  ),
  background_x = elem_list_rect(
    fill = KO_strip_text_cols
  )
)

# custom legend
steps <- c("NO3->NO2", "NO2->NO", "NO->N2O", "N2O->N2")

reduction_colors <- c(
  "NO3->NO2" = "grey60",
  "NO2->NO"  = "grey80",
  "NO->N2O"  = "#aca193",
  "N2O->N2"  = "grey90"
)

reduction_labels <- c(
  "NO3->NO2" = expression(NO[3]^'-' ~ "\u2192" ~ NO[2]^'-'),
  "NO2->NO"  = expression(NO[2]^'-' ~ "\u2192" ~ NO),
  "NO->N2O"  = expression(NO ~ "\u2192" ~ N[2]*O),
  "N2O->N2"  = expression(N[2]*O ~ "\u2192" ~ N[2])
)

legend_df <- data.frame(Step = factor(steps, levels = steps), y = seq_along(steps))
```

*FINAL MAIN: Figure 5*
```{r, fig.width=8, fig.height=10}
# stats
stat_test <- DN_KO_rel_16 %>%
  group_by(`function`) %>%
  wilcox_test(RelAbund ~ SampleType) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_xy_position(x = "SampleType")

stat_test <- stat_test %>% filter(p.adj < 0.05)

stat_test <- stat_test |>
  mutate(y.position = 0.00028)

stat_test$symbol <- "***"

# legend plot
legend_plot <- ggplot(legend_df, aes(x = 1, y = Step, fill = Step)) +
  geom_tile(color = "black", width = 0.4, height = 0.9) +
  scale_fill_manual(
    values = reduction_colors,
    labels = reduction_labels,
    name = "Denitrification Step"
  ) +
  theme_void() +
  theme(
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  )

legend_grob <- cowplot::get_legend(legend_plot)

# plot -------------------------------------------------------------------------------------------------------------------
fig5 <- DN_KO_rel_16 |>
  ggplot(aes(x = SampleType, y = RelAbund, fill = SampleType)) +
  geom_boxplot() +
  facet_wrap2(~ `function`, 
             labeller = as_labeller(KO_strip_text_labs),
             strip = strip) +
  theme_classic() +
  labs(
 #   title = "Rel Abund of denitrification KO - 16HRT",
    x = "Sample Type",
    y = "Relative Abundance"
  ) +
  scale_fill_manual(values = c(CC = "#ffa52d", WC = "darkseagreen"), name = "Media Type") +
  stat_pvalue_manual(
    stat_test,
    label = "symbol",
    tip.length = 0.01,
    hide.ns = TRUE # or FALSE to show all
  ) +
  theme(
    strip.text = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 15),
    legend.position = "bottom",
    legend.title = element_text(size = 15)
  )

# plot w/ extra legend
fig5_final <- fig5 + patchwork::wrap_elements(legend_grob) +
  plot_layout(ncol = 2, widths = c(10, 3))

#ggsave("../raw_figs_301/dnbr_fig5_picrust2.png", plot = fig5_final, width = 9, height = 10, units = "in", dpi = 300)

ggsave("../present_figs/fig5_picrust2.png", plot = fig5_final, width = 8, height = 6, units = "in", dpi = 300)
```

*NSTI plot - extra - not saved*
```{r}
nsti_317 <- read_tsv("../data/picrust2/marker_predicted_and_nsti_317.tsv")

# 317 bar plot
nsti_317 |>
  ggplot(aes(x = sequence, y = metadata_NSTI)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  labs(title = "NSTI Scores of 317 se-ASVs", x = "Each bar = Unique se-ASV (317)", y = "NSTI score") +
  scale_y_continuous(breaks = c(0.5, 1, 2), labels = c("good (0.5)", "moderate (1)", "bad (2)")) +
  theme(axis.text.x = element_blank()) +
  geom_hline(aes(yintercept = 0.5), color = "#808080", linetype = "dashed", size = 0.75) +
  geom_hline(aes(yintercept = 1), color = "#808080", linetype = "dashed", size = 0.75) +
  geom_hline(aes(yintercept = 2), color = "#cc3700", linetype = "dashed", size = 0.75) 
```

*Figure S6 A - 4-h HRT*
```{r, fig.width=8, fig.height=10}
# stats
stat_test <- DN_KO_rel_4 %>%
  group_by(`function`) %>%
  wilcox_test(RelAbund ~ SampleType) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_xy_position(x = "SampleType")

stat_test <- stat_test %>% filter(p.adj < 0.05)

stat_test <- stat_test |>
  mutate(y.position = 0.00028)

stat_test$symbol <- "***"

# legend plot
legend_plot <- ggplot(legend_df, aes(x = 1, y = Step, fill = Step)) +
  geom_tile(color = "black", width = 0.4, height = 0.9) +
  scale_fill_manual(
    values = reduction_colors,
    labels = reduction_labels,
    name = "Denitrification Step"
  ) +
  theme_void() +
  theme(
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  )

legend_grob <- cowplot::get_legend(legend_plot)

# plot
sfig6a <- DN_KO_rel_4 |>
  ggplot(aes(x = SampleType, y = RelAbund, fill = SampleType)) +
  geom_boxplot() +
  facet_wrap2(~ `function`, 
             labeller = as_labeller(KO_strip_text_labs),
             strip = strip) +
  theme_classic() +
  labs(
    title = "4-h HRT",
    x = "Media Type",
    y = "Relative Abundance"
  ) +
  scale_fill_manual(values = c(CC = "#ffa52d", WC = "darkseagreen"), name = "Media Type") +
  stat_pvalue_manual(
    stat_test,
    label = "symbol",
    tip.length = 0.01,
    hide.ns = TRUE # or FALSE to show all
  ) +
  theme(
    strip.text = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 15),
    legend.position = "bottom",
    title = element_text(size = 15)
  )
```

*Figure S6 B - 8-h HRT*
```{r, fig.width=6, fig.height=10}
# stats
stat_test <- DN_KO_rel_8 %>%
  group_by(`function`) %>%
  wilcox_test(RelAbund ~ SampleType) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_xy_position(x = "SampleType")

stat_test <- stat_test %>% filter(p.adj < 0.05)

stat_test <- stat_test |>
  mutate(y.position = 0.00028)

stat_test$symbol <- "***"

# legend plot
legend_plot <- ggplot(legend_df, aes(x = 1, y = Step, fill = Step)) +
  geom_tile(color = "black", width = 0.4, height = 0.9) +
  scale_fill_manual(
    values = reduction_colors,
    labels = reduction_labels,
    name = "Denitrification Step"
  ) +
  theme_void() +
  theme(
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  )

legend_grob <- cowplot::get_legend(legend_plot)

# plot
sfig6b <- DN_KO_rel_8 |>
  ggplot(aes(x = SampleType, y = RelAbund, fill = SampleType)) +
  geom_boxplot() +
  facet_wrap2(~ `function`, 
             labeller = as_labeller(KO_strip_text_labs),
             strip = strip) +
  theme_classic() +
  labs(
    title = "8-h HRT",
    x = "Media Type",
    y = "Relative Abundance"
  ) +
  scale_fill_manual(values = c(CC = "#ffa52d", WC = "darkseagreen"), name = "Media Type") +
  stat_pvalue_manual(
    stat_test,
    label = "symbol",
    tip.length = 0.01,
    hide.ns = TRUE # or FALSE to show all
  ) +
  theme(
    strip.text = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 15),
    legend.position = "bottom",
    title = element_text(size = 15)
  )
```

*FINAL SUPP: Figure S6*
```{r, fig.width=14, fig.height=8}
sfig6 <- sfig6a + plot_spacer() + sfig6b + plot_spacer() + patchwork::wrap_elements(legend_grob) +
  plot_layout(ncol = 5, widths = c(5, 0.1, 5, 0.1, 2))

sfig6

#ggsave("../raw_figs_301/dnbr_sfig6_48_picrust2.png", plot = sfig6, width = 14, height = 8, units = "in", dpi = 300)
#ggsave("../present_figs/dnbr_sfig6_48_picrust2.png", plot = sfig6, width = 14, height = 6, units = "in", dpi = 300)
```

*################################################################################################################*
*################################################################################################################*
#-----------------------------------------
### 7) Alpha Diversity - (FigS3, Table S2)
# Media samples only, all HRTs
#-----------------------------------------

*Prep*
```{r}
# calculate alpha diversity
alpha_df <- estimate_richness(ps_media, measures = c("Shannon", "Chao1", "Simpson"))
alpha_df$SampleID <- rownames(alpha_df)
alpha_df$SampleID <- sub("^X", "", alpha_df$SampleID)

# add sample info
alpha_df$SampleType <- sapply(alpha_df$SampleID, extract_sample_type)
alpha_df$Bioreactor <- sapply(alpha_df$SampleID, extract_bioreactor_id)
alpha_df$HRT <- sapply(alpha_df$SampleID, extract_HRT)
alpha_df$SampleSource <- sapply(alpha_df$SampleID, extract_sample_source)

# converting cols to factors
columns_to_factor <- c("SampleType", "HRT", "SampleSource", "Bioreactor")

alpha_df <- alpha_df |>
  mutate_at(vars(all_of(columns_to_factor)), ~ as.factor(as.character(.)))

## ordering HRT
HRT_order <- c("4", "8", "16")

alpha_df <- alpha_df |>
  mutate(HRT = factor(HRT, levels = HRT_order))

# numeric means - alpha diversity by sample type & HRT
alpha_df |>
  group_by(SampleType, HRT) |>
  summarise(mean_shannon = mean(Shannon, na.rm = TRUE),
            shan_std = sd(Shannon, na.rm = TRUE),
            mean_chao = mean(Chao1, na.rm = TRUE),
            chao_std = sd(Chao1, na.rm = TRUE),
            mean_simpson = mean(Simpson, na.rm = TRUE),
            simpson_std = sd(Simpson, na.rm = TRUE))
```

*Figure S3 A-C & Stats*
```{r, fig.width=4, fig.height=6}
#----------------
# Plot - Shannon
#----------------
sfig3a <- alpha_df |>
ggplot(aes(x = Bioreactor, y = Shannon)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ HRT) +
  labs(
#    title = "Shannon Diversity of CC and WC bioreactors",
    x = "Bioreactor", 
    y = "Shannon Index") +
  scale_fill_manual(
    labels = c("corncob" = "CC", "woodchip" = "WC"),
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13)) +
  scale_x_discrete(labels = c("corncob" = "CC", 
                              "woodchip" = "WC"))

#----------------
# Plot - Chao1
#----------------
sfig3b <- alpha_df |>
ggplot(aes(x = Bioreactor, y = Chao1)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ HRT) +
  labs(
#    title = "Chao1 Diversity of CC and WC bioreactors",
    x = "Bioreactor", 
    y = "Chao1 Index") +
  scale_fill_manual(
    labels = c("corncob" = "CC", "woodchip" = "WC"),
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13)) +
  scale_x_discrete(labels = c("corncob" = "CC", 
                              "woodchip" = "WC"))

#----------------
# Plot - Simpson
#----------------
sfig3c <- alpha_df |>
ggplot(aes(x = Bioreactor, y = Simpson)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ HRT) +
  labs(
#    title = "Simpson Diversity of CC and WC bioreactors",
    x = "Bioreactor", 
    y = "Simpson Index") +
  scale_fill_manual(
    labels = c("corncob" = "CC", "woodchip" = "WC"),
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13)) +
  scale_x_discrete(labels = c("corncob" = "CC", 
                              "woodchip" = "WC"))
```
```{r}
#-----------------
# Stats - Shannon 
#-----------------
alpha_df |>
  group_by(HRT) |>
  kruskal_test(Shannon ~ SampleType)

# Post Hoc 
dunn_alpha_shan <- alpha_df |>
  group_by(HRT) |>
  dunn_test(Shannon ~ Bioreactor, p.adjust.method = "bonferroni")

dunn_alpha_shan_sig <- dunn_alpha_shan |>
  filter(p.adj.signif != "ns")

#-----------------
# Stats - Chao1 
#-----------------
alpha_df |>
  group_by(HRT) |>
  kruskal_test(Chao1 ~ SampleType) 

# Post Hoc 
dunn_alpha_chao <- alpha_df |>
  group_by(HRT) |>
  dunn_test(Chao1 ~ Bioreactor, p.adjust.method = "bonferroni")

dunn_alpha_chao_sig <- dunn_alpha_chao |>
  filter(p.adj.signif != "ns")

#-----------------
# Stats - Simpson 
#-----------------

# Stats for sample type 
alpha_df |>
  group_by(HRT) |>
  kruskal_test(Simpson ~ SampleType) 

# Post Hoc 
dunn_alpha_simpson <- alpha_df |>
  group_by(HRT) |>
  dunn_test(Simpson ~ Bioreactor, p.adjust.method = "bonferroni")

dunn_alpha_simpson_sig <- dunn_alpha_simpson |>
  filter(p.adj.signif != "ns")
```

*Presentation fig?*
```{r, fig.width=12, fig.height=6}
#----------------
# Plot - Shannon
#----------------
pfig3a <- alpha_df |>
ggplot(aes(x = HRT, y = Shannon)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ SampleType) +
  labs(
#    title = "Shannon Diversity of CC and WC bioreactors",
    x = "HRT", 
    y = "Shannon Index") +
  scale_fill_manual(
    labels = c("corncob" = "CC", "woodchip" = "WC"),
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13)) +
  scale_x_discrete(labels = c("corncob" = "CC", 
                              "woodchip" = "WC"))

#----------------
# Plot - Chao1
#----------------
pfig3b <- alpha_df |>
ggplot(aes(x = HRT, y = Chao1)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ SampleType) +
  labs(
#    title = "Chao1 Diversity of CC and WC bioreactors",
    x = "HRT", 
    y = "Chao1 Index") +
  scale_fill_manual(
    labels = c("corncob" = "CC", "woodchip" = "WC"),
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 13),
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13)) +
  scale_x_discrete(labels = c("corncob" = "CC", 
                              "woodchip" = "WC"))


pfig3 <- pfig3a + plot_spacer() + pfig3b +
  plot_layout(widths = c(2, 0.1, 2), ncol = 3) 

# save plot
#ggsave("../present_figs/sfig3_alpha.png", plot = pfig3, width = 8, height = 6, units = "in", dpi = 300)

#-----------------
# Stats - Shannon 
#-----------------


# Post Hoc 


#-----------------
# Stats - Chao1 
#-----------------


```


*FINAL SUPP: Figure S3*
```{r, fig.width=30, fig.height=10}}
sfig3 <- sfig3a + plot_spacer() + sfig3b + plot_spacer() + sfig3c +
  plot_layout(widths = c(2, 0.1, 2, 0.1, 2), ncol = 5) 

# save plot
#ggsave("../raw_figs_301/dnbr_sfig3_alpha.png", plot = sfig3, width = 10, height = 6, units = "in", dpi = 300)
```

*Rarefied Analysis - media only - not saved*
```{r}
# calculate alpha diversity - ps_m_rare in 2) NMDS rarefied analysis section
alpha_df <- estimate_richness(ps_m_rare, measures = c("Shannon", "Chao1", "Simpson"))
alpha_df$SampleID <- rownames(alpha_df)
alpha_df$SampleID <- sub("^X", "", alpha_df$SampleID)

# add sample info
alpha_df$SampleType <- sapply(alpha_df$SampleID, extract_sample_type)
alpha_df$Bioreactor <- sapply(alpha_df$SampleID, extract_bioreactor_id)
alpha_df$HRT <- sapply(alpha_df$SampleID, extract_HRT)
alpha_df$SampleSource <- sapply(alpha_df$SampleID, extract_sample_source)

# converting cols to factors
columns_to_factor <- c("SampleType", "HRT", "SampleSource", "Bioreactor")

alpha_df <- alpha_df |>
  mutate_at(vars(all_of(columns_to_factor)), ~ as.factor(as.character(.)))

## ordering HRT
HRT_order <- c("4", "8", "16")

alpha_df <- alpha_df |>
  mutate(HRT = factor(HRT, levels = HRT_order))
```
```{r, fig.width=4, fig.height=6}
#----------------
# Plot - Shannon
#----------------
alpha_df |>
ggplot(aes(x = Bioreactor, y = Shannon)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ HRT) +
  labs(
#    title = "Shannon Diversity of CC and WC bioreactors",
    x = "Bioreactor", 
    y = "Shannon Index") +
  scale_fill_manual(
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Sample Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 15),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15))

#----------------
# Plot - Chao1
#----------------
alpha_df |>
ggplot(aes(x = Bioreactor, y = Chao1)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ HRT) +
  labs(
#    title = "Chao1 Diversity of CC and WC bioreactors",
    x = "Bioreactor", 
    y = "Chao1 Index") +
  scale_fill_manual(
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Sample Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 15),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15))

#----------------
# Plot - Simpson
#----------------
alpha_df |>
ggplot(aes(x = Bioreactor, y = Simpson)) +
  geom_boxplot(aes(fill = SampleType)) +
#  geom_jitter(width = 0.2, size = 2, alpha = 0.4, aes(shape = SampleSource)) +
  facet_wrap(~ HRT) +
  labs(
#    title = "Simpson Diversity of CC and WC bioreactors",
    x = "Bioreactor", 
    y = "Simpson Index") +
  scale_fill_manual(
    values = c("corncob" = "#ffa52d", "woodchip" = "darkseagreen"),
    name = "Sample Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    scale_shape_manual(values = c("media" = 5, "water" = 16)) +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 15),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15))
```

*################################################################################################################*
*################################################################################################################*
#-------------------------------
### 8) TOC
# Media samples only, all HRTs
#-------------------------------

*TOC boxplots & means & stats*
```{r, fig.width=7, fig.height=6}

# ----------- REMOVING OUTLIERS -----------

nre_data$SampleID <- ifelse(
  nre_data$HRT == 4,
  paste0("4", paste(nre_data$sample_type, nre_data$bioreactor_ID, nre_data$influent_id, sep = "_")),
  paste(nre_data$sample_type, nre_data$bioreactor_ID, nre_data$influent_id, sep = "_")
)

nre_data_removed_outliers <- nre_data |>
  filter(!SampleID %in% c("CC_1_28", "CC_2_28", "CC_3_28", "WC_4_7"))

#------------------------------------------------
# Plot - TOC comparison by bioreactor rep (1-6)
#------------------------------------------------
toc_br <- nre_data_removed_outliers |>
  ggplot(aes(bioreactor_ID, TOC_conc, fill = sample_type)) +
  geom_boxplot() +
  facet_wrap(~HRT) +
  labs(
    x = "Bioreactor ID",
    y = "TOC (mg C/L)"
  ) +
  theme_bw() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

#-------------------------------------------------
# Plot - TOC comparison by sample type (CC vs WC)
#-------------------------------------------------
toc_st <- nre_data_removed_outliers |>
  ggplot(aes(sample_type, TOC_conc, fill = sample_type)) +
  geom_boxplot() +
  facet_wrap(~HRT) +
  labs(
    x = "Media Type",
    y = "TOC (mg C/L)"
  ) +
  theme_bw() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

# numeric means
toc_means_df <- nre_data |>
  group_by(sample_type, HRT) |>
  summarise(mean_conc= mean(TOC_conc, na.rm = TRUE),
            std = sd(TOC_conc, na.rm = TRUE))

#-------------------------------------------------
# Plot - TOC comparison by sample type (CC vs WC)
#-------------------------------------------------
toc_st_facet <- nre_data_removed_outliers |>
  ggplot(aes(HRT, TOC_conc, fill = sample_type)) +
  geom_boxplot() +
  facet_wrap(~sample_type) +
  labs(
    x = "Media Type",
    y = "TOC (mg C/L)"
  ) +
  theme_bw() +
  scale_fill_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type") +
  scale_y_continuous(limits = c(0, 110)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 15),
        strip.background = element_rect(fill = "grey85"),
        panel.grid = element_blank(),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 13),
        legend.text = element_text(size = 13))

# full model comparison [KRUSKAL] - p < 0.05 = signif diff in TOC between HRT (CC & WC: p < 0.05 by a lot (e-16))
nre_data_removed_outliers |>
  group_by(sample_type) |>
  kruskal_test(TOC_conc ~ HRT) 

# pairwise comparison [DUNN] 
toc_dunn_st <- nre_data_removed_outliers |>
  group_by(sample_type) |>
  dunn_test(TOC_conc ~ HRT, p.adjust.method = "bonferroni")

toc_dunn_st_sig <- nre_dunn_br |>
  filter(p.adj.signif != "ns")

#------------------------------------
# Stats - NRE separated by CC vs WC
#------------------------------------

# full model comparison [KRUSKAL] - p < 0.05 = signif diff in removal efficiency between sample type
nre_data_removed_outliers |>
  group_by(HRT) |>
  kruskal_test(TOC_conc ~ sample_type)

#------------------------------------------------
# Stats - NRE separated by bioreactor rep (1-6)
#------------------------------------------------
# pairwise comparison [DUNN]
nre_dunn_ST <- nre_data_removed_outliers |>
  group_by(HRT) |>
  dunn_test(removal_efficiency ~ bioreactor_ID, p.adjust.method = "bonferroni")

nre_dunn_ST_sig <- nre_dunn_ST |>
  filter(p.adj.signif != "ns")

```

*TOC & NRE correlation*
```{r, fig.width=15, fig.height=8}

# calculating Spearman's correlation - ALL DATA
toc_nre_corr_stats <- nre_data |>
  group_by(HRT, sample_type) |>
  dplyr::summarize(
    rho = cor(TOC_conc, removal_efficiency, method = "spearman", use = "complete.obs"),
    p   = cor.test(TOC_conc, removal_efficiency, method = "spearman")$p.value,
    .groups = "drop") |>
  mutate(
     sig = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ ""
  ))

#--------------------------------------------------
# Plot - TOC vs NRE in HRT for CC & WC - ALL DATA
#--------------------------------------------------
toc_nre_corr_p <- nre_data |>
  ggplot(aes(x = TOC_conc, y = removal_efficiency, group = sample_type, colour = sample_type)) +
  geom_point(size = 3) +
  geom_smooth(aes(group = sample_type), method = "lm", se = FALSE, linetype = "dashed") +
  stat_cor(method = "spearman", 
           aes(label = paste(factor(..group.., labels = c("CC", "WC")), ": ", ..r.label.., ..p.label.., sep = " ")),
           output.type = "text",
           label.x.npc = 0.4, label.y.npc = 0.3,
           size = 5, 
           color = "black") +
  labs(
    x = "Total organic carbon (mg C/L)",
    y = "Nitrate Removal Efficiency (%)",
    color = "Media Type"
  ) +
  theme_classic() +
  facet_wrap(~ HRT, scales = "free_y") +
  scale_color_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type"
  ) +
    theme(legend.position = "bottom",
        axis.text = element_text(size = 20), 
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        axis.title.y = element_text(margin = margin(r = 10)))

# ----------- REMOVING OUTLIERS -----------

nre_data$SampleID <- ifelse(
  nre_data$HRT == 4,
  paste0("4", paste(nre_data$sample_type, nre_data$bioreactor_ID, nre_data$influent_id, sep = "_")),
  paste(nre_data$sample_type, nre_data$bioreactor_ID, nre_data$influent_id, sep = "_")
)

# calculating Spearman's correlation - ALL DATA
nre_data_removed_outliers <- nre_data |>
  filter(!SampleID %in% c("CC_1_28", "CC_2_28", "CC_3_28", "WC_4_7"))

toc_nre_corr_stats_RO <- nre_data_removed_outliers |>
  group_by(HRT, sample_type) |>
  dplyr::summarize(
    rho = cor(TOC_conc, removal_efficiency, method = "spearman", use = "complete.obs"),
    p   = cor.test(TOC_conc, removal_efficiency, method = "spearman")$p.value,
    .groups = "drop") |>
  mutate(
     sig = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ ""
  ))

#---------------------------------------------------------
# Plot - TOC vs NRE in HRT for CC & WC - REMOVED OUTLIERS
#---------------------------------------------------------
toc_nre_corr_RO_p <- nre_data_removed_outliers |>
  ggplot(aes(x = TOC_conc, y = removal_efficiency, group = sample_type, colour = sample_type)) +
  geom_point(size = 3) +
  geom_smooth(aes(group = sample_type), method = "lm", se = FALSE, linetype = "dashed") +
  stat_cor(method = "spearman", 
           aes(label = paste(factor(..group.., labels = c("CC", "WC")), ": ", ..r.label.., ..p.label.., sep = " ")),
           output.type = "text",
           label.x.npc = 0.4, label.y.npc = 0.3,
           size = 5, 
           color = "black") +
  labs(
    x = "Total organic carbon (mg C/L)",
    y = "Nitrate Removal Efficiency (%)",
    color = "Media Type"
  ) +
  theme_classic() +
  facet_wrap(~ HRT, scales = "free_y") +
  scale_color_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type"
  ) +
    theme(legend.position = "bottom",
        axis.text = element_text(size = 20), 
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        axis.title.y = element_text(margin = margin(r = 10)))

#---------------------------------------------------------
# Plot - TOC vs NRE in HRT for CC & WC - REMOVED OUTLIERS
#---------------------------------------------------------
toc_nre_corr_RO_allHRT_p <- nre_data_removed_outliers |>
  ggplot(aes(x = TOC_conc, y = removal_efficiency, group = sample_type, colour = sample_type)) +
  geom_point(size = 3) +
  geom_smooth(aes(group = sample_type), method = "lm", se = FALSE, linetype = "dashed") +
  stat_cor(method = "spearman", 
           aes(label = paste(factor(..group.., labels = c("CC", "WC")), ": ", ..r.label.., ..p.label.., sep = " ")),
           output.type = "text",
           label.x.npc = 0.4, label.y.npc = 0.3,
           size = 5, 
           color = "black") +
  labs(
    x = "Total organic carbon (mg C/L)",
    y = "Nitrate Removal Efficiency (%)",
    color = "Media Type"
  ) +
  theme_classic() +
#  facet_wrap(~ HRT, scales = "free_y") +
  scale_color_manual(
    values = c("CC" = "#ffa52d", "WC" = "darkseagreen"),
    name = "Media Type"
  ) +
    theme(legend.position = "bottom",
        axis.text = element_text(size = 20), 
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        axis.title.y = element_text(margin = margin(r = 10)))
```

*TOC & se-ASVs* <- Must have objects from 5) Correlation section read in
```{r, fig.width=12, fig.height=8}
# plot - 71 CC se-ASVs & TOC -------------------------------------------------------------------------------------------------------------------
CC_71_corr_toc <- nre_CC_pso_df |>
  ggplot(aes(x = TOC_conc, y = SumTaxa, shape = HRT, colour = SampleType)) +
  geom_point(size = 3) +
  stat_cor(method = "spearman", 
           aes(label = paste0("rho = ", ..r.., ", p = ", formatC(..p.., format = "e", digits = 2)), group = 1),
           output.type = "text",
           label.x.npc = 0.5, label.y.npc = 0.4,
           size = 5, 
           color = "black") +
  labs(
    x = "Total Organic Carbon (mg C/L)",
    y = "Relative Abundance of CC se-ASVs (71)",
    color = "Sample Type",
#    title = "CC se-ASVs only"
  ) +
  theme_classic() +
  scale_color_manual(
    values = c("CC" = "black"), name = "Media Type"
  ) + 
  scale_shape_manual(
    values = c(
      "4" = 18, 
      "8" = 17,
      "16" = 16
    )
  ) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 20), 
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title.y = element_text(margin = margin(r = 10)))

# checking small p-value <- p = 2.2e-16, rho = 0.8637639 
cor.test(
  nre_CC_pso_df$TOC_conc,
  nre_CC_pso_df$SumTaxa,
  method = "spearman",
  exact  = FALSE   
)

# plot - 246 WC se-ASVs & TOC -------------------------------------------------------------------------------------------------------------------
WC_246_corr_toc <- nre_WC_pso_df |>
  ggplot(aes(x = TOC_conc, y = SumTaxa, shape = HRT, colour = SampleType)) +
  geom_point(size = 3) +
  stat_cor(method = "spearman", 
           aes(label = paste0("rho = ", ..r.., ", p = ", formatC(..p.., format = "e", digits = 2)), group = 1),
           output.type = "text",
           label.x.npc = 0.5, label.y.npc = 0.4,
           size = 5, 
           color = "black") +
  labs(
    x = "Total Organic Carbon (mg C/L)",
    y = "Relative Abundance of WC se-ASVs (246)",
    color = "Sample Type",
#    title = "WC se-ASVs only"
  ) +
  theme_classic() +
  scale_color_manual(
    values = c("WC" = "black"), name = "Media Type"
  ) + 
  scale_shape_manual(
    values = c(
      "4" = 18, 
      "8" = 17,
      "16" = 16
    )
  ) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 20), 
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title.y = element_text(margin = margin(r = 10)))


# checking small p-value <- p = 1.657e-05, rho = 0.5837188 
cor.test(
  nre_WC_pso_df$TOC_conc,
  nre_WC_pso_df$SumTaxa,
  method = "spearman",
  exact  = FALSE  
)

CC_71_corr_toc + plot_spacer() + WC_246_corr_toc + 
  plot_layout(widths = c(2, 0.1, 2), ncol = 3) 
```

*################################################################################################################*
*################################################################################################################*
#-------------------------------------------
### 9) Taxonomy 
# 317 se-ASVs, media samples only, all HRTs
#-------------------------------------------

#---------------- BAR ----------------

*CC - se-asv Phyla by Bioreactor*
```{r}
ps_media_CC <- subset_samples(ps_media, SampleType == "corncob")

#---------------------------------------------------
# Calculating relative abundance HRT and Bioreactor
#---------------------------------------------------
ps_media_CC_rel_df <- ps_media_CC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Phylum), Phylum != "NA") |>
  group_by(Bioreactor, HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_CC_rel_df$Phylum_ID <- ifelse(ps_media_CC_rel_df$OTU %in% CC_71_se_asvs, as.character(ps_media_CC_rel_df$Phylum), "Other") 

# # check
# check_df <- ps_media_CC_rel_df |>
#   filter(OTU %in% CC_71_se_asvs)

# ordering Phyla
unique(ps_media_CC_rel_df$Phylum_ID)

se_asv_phylum_map <- c(
  "Proteobacteria" = "#1F77B4", 
  "Actinobacteriota" = "#2CA02C",
  "Cyanobacteria" = "#9467BD",
  "Euryarchaeota" = "#8C564B",
  "Bacteroidota" = "#393B79",
  "Acidobacteriota" = "#637939",
  "Desulfobacterota" = "#7B4173",
  "Planctomycetota" = "#AD494A",
  "Myxococcota" = "#BEAED4",
  "Verrucomicrobiota" = "#A6CEE3",
  "Firmicutes" = "#B5CF6B",
  "Chloroflexi" = "#E7BA52",
  "Other" = "#999999"
)

# Phylum_ID to factor for ordering
ps_media_CC_rel_df$Phylum_ID <- factor(ps_media_CC_rel_df$Phylum_ID,
                                     levels = names(se_asv_phylum_map))
```
```{r, fig.height=8, fig.width=10}
# plot
cc_se_asv_phyla_br <- ggplot(ps_media_CC_rel_df, aes(x = Bioreactor, y = relative_abund, fill = Phylum_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  facet_wrap(~HRT) +
  scale_fill_manual(name = "CC se-ASV Phyla", values = se_asv_phylum_map) +
  labs(
    x = "Bioreactor", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/cc_se-asv_phyla_BR.png", plot = cc_se_asv_phyla_br, width = 8, height = 6, units = "in", dpi = 300)
```

*CC - se-asv Phyla by HRT*
```{r}
#------------------------------------
# Calculating relative abundance HRT
#------------------------------------
ps_media_CC_rel_df <- ps_media_CC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Phylum), Phylum != "NA") |>
  group_by(HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_CC_rel_df$Phylum_ID <- ifelse(ps_media_CC_rel_df$OTU %in% CC_71_se_asvs, as.character(ps_media_CC_rel_df$Phylum), "Other") 

# # check
# check_df <- ps_media_CC_rel_df |>
#   filter(OTU %in% CC_71_se_asvs)

# ordering Phyla
unique(ps_media_CC_rel_df$Phylum_ID)

se_asv_phylum_map <- c(
  "Proteobacteria" = "#1F77B4", 
  "Actinobacteriota" = "#2CA02C",
  "Cyanobacteria" = "#9467BD",
  "Euryarchaeota" = "#8C564B",
  "Bacteroidota" = "#393B79",
  "Acidobacteriota" = "#637939",
  "Desulfobacterota" = "#7B4173",
  "Planctomycetota" = "#AD494A",
  "Myxococcota" = "#BEAED4",
  "Verrucomicrobiota" = "#A6CEE3",
  "Firmicutes" = "#B5CF6B",
  "Chloroflexi" = "#E7BA52",
  "Other" = "#999999"
)

# Phylum_ID to factor for ordering
ps_media_CC_rel_df$Phylum_ID <- factor(ps_media_CC_rel_df$Phylum_ID,
                                     levels = names(se_asv_phylum_map))
```
```{r, fig.height=8, fig.width=6}
# plot
cc_se_asv_phyla_br <- ggplot(ps_media_CC_rel_df, aes(x = HRT, y = relative_abund, fill = Phylum_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  scale_fill_manual(name = "CC se-ASV Phyla", values = se_asv_phylum_map) +
  labs(
    x = "HRT", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/cc_se-asv_phyla_HRT.png", plot = cc_se_asv_phyla_br, width = 8, height = 6, units = "in", dpi = 300)
```

*WC - se-asv Phyla by Bioreactor*
```{r}
ps_media_WC <- subset_samples(ps_media, SampleType == "woodchip")

#---------------------------------------------------
# Calculating relative abundance HRT and Bioreactor
#---------------------------------------------------
ps_media_WC_rel_df <- ps_media_WC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Phylum), Phylum != "NA") |>
  group_by(Bioreactor, HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_WC_rel_df$Phylum_ID <- ifelse(ps_media_WC_rel_df$OTU %in% WC_246_se_asvs, as.character(ps_media_WC_rel_df$Phylum), "Other") 

# # check
# check_df <- ps_media_WC_rel_df |>
#   filter(OTU %in% WC_246_se_asvs)

# ordering Phyla
unique(ps_media_WC_rel_df$Phylum_ID)

se_asv_phylum_map <- c(
  "Proteobacteria" = "#1F77B4", 
  "Actinobacteriota" = "#2CA02C",
  "Cyanobacteria" = "#9467BD",
  "Bacteroidota" = "#393B79",
  "Acidobacteriota" = "#637939",
  "Desulfobacterota" = "#7B4173",
  "Planctomycetota" = "#AD494A",
  "Myxococcota" = "#BEAED4",
  "Verrucomicrobiota" = "#A6CEE3",
  "Firmicutes" = "#B5CF6B",
  "Chloroflexi" = "#E7BA52",
  "Halobacterota" = "#CEDB9C",
  "Thermoplasmatota" = "#CE6DBD",
  "Patescibacteria" = "#1B9E77",
  "Spirochaetota" = "#FF7F0E",
  "Armatimonadota" = "#17BECF",
  "Bdellovibrionota" = "#D62728",
  "Gemmatimonadota" = "#E6AB02",
  "Other" = "#999999"
)

# Phylum_ID to factor for ordering
ps_media_WC_rel_df$Phylum_ID <- factor(ps_media_WC_rel_df$Phylum_ID,
                                     levels = names(se_asv_phylum_map))
```
```{r, fig.height=8, fig.width=10}
# plot
WC_se_asv_phyla_br <- ggplot(ps_media_WC_rel_df, aes(x = Bioreactor, y = relative_abund, fill = Phylum_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  facet_wrap(~HRT) +
  scale_fill_manual(name = "WC se-ASV Phyla", values = se_asv_phylum_map) +
  labs(
    x = "Bioreactor", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/WC_se-asv_phyla_BR.png", plot = WC_se_asv_phyla_br, width = 8, height = 6, units = "in", dpi = 300)
```

*WC - se-asv Phyla by HRT*
```{r}
#------------------------------------
# Calculating relative abundance HRT
#------------------------------------
ps_media_WC_rel_df <- ps_media_WC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Phylum), Phylum != "NA") |>
  group_by(HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_WC_rel_df$Phylum_ID <- ifelse(ps_media_WC_rel_df$OTU %in% WC_246_se_asvs, as.character(ps_media_WC_rel_df$Phylum), "Other") 

# # check
# check_df <- ps_media_WC_rel_df |>
#   filter(OTU %in% WC_246_se_asvs)

# ordering Phyla
unique(ps_media_WC_rel_df$Phylum_ID)

se_asv_phylum_map <- c(
  "Proteobacteria" = "#1F77B4", 
  "Actinobacteriota" = "#2CA02C",
  "Cyanobacteria" = "#9467BD",
  "Bacteroidota" = "#393B79",
  "Acidobacteriota" = "#637939",
  "Desulfobacterota" = "#7B4173",
  "Planctomycetota" = "#AD494A",
  "Myxococcota" = "#BEAED4",
  "Verrucomicrobiota" = "#A6CEE3",
  "Firmicutes" = "#B5CF6B",
  "Chloroflexi" = "#E7BA52",
  "Halobacterota" = "#CEDB9C",
  "Thermoplasmatota" = "#CE6DBD",
  "Patescibacteria" = "#1B9E77",
  "Spirochaetota" = "#FF7F0E",
  "Armatimonadota" = "#17BECF",
  "Bdellovibrionota" = "#D62728",
  "Gemmatimonadota" = "#E6AB02",
  "Other" = "#999999"
)

# Phylum_ID to factor for ordering
ps_media_WC_rel_df$Phylum_ID <- factor(ps_media_WC_rel_df$Phylum_ID,
                                     levels = names(se_asv_phylum_map))
```
```{r, fig.height=8, fig.width=6}
# plot
WC_se_asv_phyla_br <- ggplot(ps_media_WC_rel_df, aes(x = HRT, y = relative_abund, fill = Phylum_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  scale_fill_manual(name = "WC se-ASV Phyla", values = se_asv_phylum_map) +
  labs(
    x = "HRT", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/WC_se-asv_phyla_HRT.png", plot = WC_se_asv_phyla_br, width = 8, height = 6, units = "in", dpi = 300)
```

*CC - se-asv Genera by Bioreactor*
```{r}
ps_media_CC <- subset_samples(ps_media, SampleType == "corncob")

#---------------------------------------------------
# Calculating relative abundance HRT and Bioreactor
#---------------------------------------------------
ps_media_CC_rel_df <- ps_media_CC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Genus), Genus != "NA") |>
  group_by(Bioreactor, HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_CC_rel_df$Genus_ID <- ifelse(ps_media_CC_rel_df$OTU %in% CC_71_se_asvs, as.character(ps_media_CC_rel_df$Genus), "Other") 

# # check
# check_df <- ps_media_CC_rel_df |>
#   filter(OTU %in% CC_71_se_asvs)

# ordering Phyla
unique(ps_media_CC_rel_df$Genus_ID)

se_asv_genus_cols <- colorRampPalette(brewer.pal(12, "Paired"))(50)
se_asv_genus_cols

se_asv_genus_map <- c(
  "Xanthobacter" = "#1F77B4", 
  "Pseudorhodoplanes" = "#2CA02C",
  "Azotobacter" = "#9467BD",
  "Burkholderia-Caballeronia-Paraburkholderia" = "#8C564B",
  "Cellulomonas" = "#393B79",
  "Azospira" = "#637939",
  "Zoogloea" = "#7B4173",
  "Pseudomonas" = "#AD494A",
  "Plot4-2H12" = "#BEAED4",
  "Methanobacterium" = "#A6CEE3",
  "Mycobacterium" = "#B5CF6B",
  "Bradyrhizobium" = "#E7BA52",
  "Nevskia" = "#1F77B4", 
  "Sphingomonas" = "#2CA02C",
  "Geothrix" = "#9467BD",
  "Desulfovibrio" = "#8C564B",
  "Mesorhizobium" = "#393B79",
  "Rhizomicrobium" = "#637939",
  "Caulobacter" = "#7B4173",
  "Pleomorphomonas" = "#AD494A",
  "Coxiella" = "#BEAED4",
  "Sediminibacterium" = "#A6CEE3",
  "Pajaroellobacter" = "#B5CF6B",
  "Undibacterium" = "#E7BA52",
  "Edaphobaculum" = "#1F77B4", 
  "Pseudolabrys" = "#2CA02C",
  "Kaistia" = "#9467BD",
  "Hyphomicrobium" = "#8C564B",
  "Rudaea" = "#393B79",
  "Devosia" = "#637939",
  "Aetherobacter" = "#7B4173",
  "Prosthecobacter" = "#AD494A",
  "Beijerinckia" = "#BEAED4",
  "Puia" = "#A6CEE3",
  "Novosphingobium" = "#B5CF6B",
  "Dyella" = "#E7BA52",
  "LD29" = "#1F77B4", 
  "Anaerolineaceae UCG-001" = "#2CA02C",
  "Bacteroides" = "#9467BD",
  "Papillibacter" = "#8C564B",
  "Sphingobium" = "#393B79",
  "Gluconacetobacter" = "#637939",
  "Tahibacter" = "#7B4173",
  "Gemmata" = "#AD494A",
  "Rhodopila" = "#BEAED4",
  "Ferruginibacter" = "#A6CEE3",
  "Thermincola" = "#B5CF6B",
  "WCHB1-32" = "#E7BA52",
  "Limnohabitans" = "#BEAED4",
  "Other" = "#999999"
)

# Genus_ID to factor for ordering
ps_media_CC_rel_df$Genus_ID <- factor(ps_media_CC_rel_df$Genus_ID,
                                     levels = names(se_asv_genus_map))
```
```{r, fig.height=8, fig.width=10}
# plot
cc_se_asv_genera_br <- ggplot(ps_media_CC_rel_df, aes(x = Bioreactor, y = relative_abund, fill = Genus_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  facet_wrap(~HRT) +
  scale_fill_manual(name = "CC se-ASV Phyla", values = se_asv_genus_map) +
  labs(
    x = "Bioreactor", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/cc_se-asv_genera_BR.png", plot = cc_se_asv_genera_br, width = 8, height = 6, units = "in", dpi = 300)
```

*CC - se-asv Genera by HRT*
```{r}
#------------------------------------
# Calculating relative abundance HRT
#------------------------------------
ps_media_CC_rel_df <- ps_media_CC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Genus), Genus != "NA") |>
  group_by(HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_CC_rel_df$Genus_ID <- ifelse(ps_media_CC_rel_df$OTU %in% CC_71_se_asvs, as.character(ps_media_CC_rel_df$Genus), "Other") 

# # check
# check_df <- ps_media_CC_rel_df |>
#   filter(OTU %in% CC_71_se_asvs)

# ordering Phyla
unique(ps_media_CC_rel_df$Genus_ID)



# Genus_ID to factor for ordering
ps_media_CC_rel_df$Genus_ID <- factor(ps_media_CC_rel_df$Genus_ID,
                                     levels = names(se_asv_Genus_map))
```
```{r, fig.height=8, fig.width=6}
# plot
cc_se_asv_phyla_br <- ggplot(ps_media_CC_rel_df, aes(x = HRT, y = relative_abund, fill = Genus_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  scale_fill_manual(name = "CC se-ASV Phyla", values = se_asv_Genus_map) +
  labs(
    x = "HRT", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/cc_se-asv_phyla_HRT.png", plot = cc_se_asv_phyla_br, width = 8, height = 6, units = "in", dpi = 300)
```

*WC - se-asv Genera by Bioreactor*
```{r}
ps_media_WC <- subset_samples(ps_media, SampleType == "woodchip")

#---------------------------------------------------
# Calculating relative abundance HRT and Bioreactor
#---------------------------------------------------
ps_media_WC_rel_df <- ps_media_WC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Genus), Genus != "NA") |>
  group_by(Bioreactor, HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_WC_rel_df$Genus_ID <- ifelse(ps_media_WC_rel_df$OTU %in% WC_246_se_asvs, as.character(ps_media_WC_rel_df$Genus), "Other") 

# # check
# check_df <- ps_media_WC_rel_df |>
#   filter(OTU %in% WC_246_se_asvs)

# ordering Phyla
unique(ps_media_WC_rel_df$Genus_ID)



# Genus_ID to factor for ordering
ps_media_WC_rel_df$Genus_ID <- factor(ps_media_WC_rel_df$Genus_ID,
                                     levels = names(se_asv_Genus_map))
```
```{r, fig.height=8, fig.width=10}
# plot
WC_se_asv_phyla_br <- ggplot(ps_media_WC_rel_df, aes(x = Bioreactor, y = relative_abund, fill = Genus_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  facet_wrap(~HRT) +
  scale_fill_manual(name = "WC se-ASV Phyla", values = se_asv_Genus_map) +
  labs(
    x = "Bioreactor", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/WC_se-asv_phyla_BR.png", plot = WC_se_asv_phyla_br, width = 8, height = 6, units = "in", dpi = 300)
```

*WC - se-asv Genera by HRT*
```{r}
#------------------------------------
# Calculating relative abundance HRT
#------------------------------------
ps_media_WC_rel_df <- ps_media_WC |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Genus), Genus != "NA") |>
  group_by(HRT) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping for phyla coloring
ps_media_WC_rel_df$Genus_ID <- ifelse(ps_media_WC_rel_df$OTU %in% WC_246_se_asvs, as.character(ps_media_WC_rel_df$Genus), "Other") 

# # check
# check_df <- ps_media_WC_rel_df |>
#   filter(OTU %in% WC_246_se_asvs)

# ordering Phyla
unique(ps_media_WC_rel_df$Genus_ID)



# Genus_ID to factor for ordering
ps_media_WC_rel_df$Genus_ID <- factor(ps_media_WC_rel_df$Genus_ID,
                                     levels = names(se_asv_Genus_map))
```
```{r, fig.height=8, fig.width=6}
# plot
WC_se_asv_phyla_br <- ggplot(ps_media_WC_rel_df, aes(x = HRT, y = relative_abund, fill = Genus_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  scale_fill_manual(name = "WC se-ASV Phyla", values = se_asv_Genus_map) +
  labs(
    x = "HRT", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15), 
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))

#ggsave("../raw_figs_301/WC_se-asv_phyla_HRT.png", plot = WC_se_asv_phyla_br, width = 8, height = 6, units = "in", dpi = 300)
```

*CC & WC - all 317 se-asv Phyla by triplicate ID*
```{r, fig.height=10, fig.width=8}
#-------------------------------------------------------------
# Calculating Phyla relative abundance by HRT & Triplicate ID 
#-------------------------------------------------------------
ps_media_rel_df <- ps_media |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Phylum), Phylum != "NA")

# making a triplicate column
ps_media_rel_df$triplicate_id <- str_replace(ps_media_rel_df$SampleID, "_\\d+_(\\d+)", "_\\1")

ps_media_rel_df <- ps_media_rel_df |>
  group_by(triplicate_id) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping <- all 317, not split by media-specific se-ASV
ps_media_rel_df$Phylum_ID <- ifelse(
  ps_media_rel_df$OTU %in% se_asv_317, 
  as.character(ps_media_rel_df$Phylum), 
  "Other"
)

# Phylum ID to factor
unique(ps_media_rel_df$Phylum_ID)

se_asv_phylum_map <- c(
  "Proteobacteria" = "#1F77B4", 
  "Actinobacteriota" = "#2CA02C",
  "Cyanobacteria" = "#9467BD",
  "Euryarchaeota" = "#8C564B",
  "Bacteroidota" = "#393B79",
  "Acidobacteriota" = "#637939",
  "Desulfobacterota" = "#7B4173",
  "Planctomycetota" = "#AD494A",
  "Myxococcota" = "#BEAED4",
  "Verrucomicrobiota" = "#A6CEE3",
  "Firmicutes" = "#B5CF6B",
  "Chloroflexi" = "#E7BA52",
  "Halobacterota" = "#CEDB9C",
  "Thermoplasmatota" = "#CE6DBD",
  "Patescibacteria" = "#1B9E77",
  "Spirochaetota" = "#FF7F0E",
  "Armatimonadota" = "#17BECF",
  "Bdellovibrionota" = "#D62728",
  "Gemmatimonadota" = "#E6AB02",
  "Other" = "#999999"
)

ps_media_rel_df$Phylum_ID <- factor(ps_media_rel_df$Phylum_ID,
                                     levels = names(se_asv_phylum_map))

ggplot(ps_media_rel_df, aes(x = triplicate_id, y = relative_abund, fill = Phylum_ID)) +
  geom_bar(stat = "identity", position = "stack", color = NA, na.rm = TRUE) +
  theme_bw() +
  facet_wrap(~ SampleType + HRT, scales = "free_x") +
  scale_fill_manual(name = "CC se-ASV Phyla", values = se_asv_phylum_map) +
  labs(
    x = "HRT", 
    y = "Relative Abundance"
  ) +
  theme(panel.grid = element_blank(),
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 15),
        axis.text.x = element_text(angle = 90),
        strip.text = element_text(size = 15), 
        axis.title = element_text(size = 15))
```

#---------------- HEATMAP ----------------

*CC & WC 317 se-asv Phyla by triplicate ID*
```{r, fig.height=6, fig.width=12}
ps_media_rel_df <- ps_media |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Phylum), Phylum != "NA")

# making a triplicate column
ps_media_rel_df$triplicate_id <- str_replace(ps_media_rel_df$SampleID, "_\\d+_(\\d+)", "_\\1")

ps_media_rel_df <- ps_media_rel_df |>
  group_by(triplicate_id) |>
  mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping <- all 317, not split by media-specific se-ASV
ps_media_rel_df$Phylum_ID <- ifelse(
  ps_media_rel_df$OTU %in% se_asv_317, 
  as.character(ps_media_rel_df$Phylum), 
  "Other"
)

ps_media_heat_df <- ps_media_rel_df |>
  group_by(
    SampleType,
    HRT,
    triplicate_id,
    Phylum_ID
  ) |>
  dplyr::summarise(
    rel_abund = sum(relative_abund, na.rm = TRUE),
    .groups = "drop"
  )

phyla_factor_order <- c("Proteobacteria", "Actinobacteriota", "Cyanobacteria", "Euryarchaeota", "Bacteroidota", "Acidobacteriota", "Desulfobacterota", "Planctomycetota",
                        "Myxococcota", "Verrucomicrobiota", "Firmicutes", "Chloroflexi", "Halobacterota", "Thermoplasmatota", "Patescibacteria", "Spirochaetota", 
                        "Armatimonadota", "Bdellovibrionota", "Gemmatimonadota", "Other")

ps_media_heat_df$Phylum_ID <- factor(
  ps_media_heat_df$Phylum_ID, 
  levels = rev(phyla_factor_order))

# bin relative abundance for plot
ps_media_heat_df <- ps_media_heat_df %>%
  mutate(
    rel_abund_bin = cut(
      rel_abund,
      breaks = c(0, 0.1, 0.2, 0.4, 0.6, 0.8, 1),
      include.lowest = TRUE,
      right = FALSE,
      labels = c(
        "0–0.1",
        "0.1–0.2",
        "0.2–0.4",
        "0.4–0.6",
        "0.6–0.8",
        "0.8–1"
      )
    )
  )

# bin color map
heatmap_colors <- c(
  "0–0.1"   = "#E0E9FF",
  "0.1–0.2" = "#ADC6FE",
  "0.2–0.4" = "#759FFD",
  "0.4–0.6" = "#2366FC",
  "0.6–0.8" = "#02329D",
  "0.8–1"   = "#011C58"
)

ps_media_heat_df <- ps_media_heat_df %>%
  mutate(
    rel_abund_bin = ifelse(
      rel_abund == 0,
      "0",
      as.character(rel_abund_bin)
    ),
    rel_abund_bin = factor(
      rel_abund_bin,
      levels = c("0", names(heatmap_colors))
    )
  )

heatmap_colors <- c(
  "0"       = "white",
  heatmap_colors
)


# optional: drop "other"
# ps_media_heat_df <- ps_media_heat_df |>
#   filter(Phylum_ID != "Other")

# ordering triplicates
unique(ps_media_heat_df$triplicate_id)

triplicate_order <- c("4CC_0", "4CC_4", "4CC_7","4CC_11", "4CC_14", "4CC_18", "4CC_21",  
                      "CC_6", "CC_10", "CC_15", "CC_20",  
                      "CC_21", "CC_28", "CC_29", "CC_30", "CC_31", "CC_32", "CC_33", "CC_34", "CC_35", "CC_36", "CC_37", "CC_38", 
                      "4WC_0", "4WC_4", "4WC_8", "4WC_11", "4WC_14", "4WC_18", "4WC_21",  
                      "WC_6", "WC_10", "WC_15", "WC_20", 
                      "WC_21", "WC_24", "WC_27", "WC_28", "WC_29", "WC_30", "WC_31")

ps_media_heat_df$triplicate_id <- factor(ps_media_heat_df$triplicate_id, levels = triplicate_order)

new_phyla_taxa_names <- c(
  "Proteobacteria" = "Pseudomonadota", 
  "Actinobacteriota" = "Actinomycetota",
  "Cyanobacteria" = "Cyanobacteria",
  "Euryarchaeota" = "Euryarchaeota",
  "Bacteroidota" = "Bacteroidota",
  "Acidobacteriota" = "Acidobacteriota",
  "Desulfobacterota" = "Desulfobacterota",
  "Planctomycetota" = "Planctomycetota",
  "Myxococcota" = "Myxococcota",
  "Verrucomicrobiota" = "Verrucomicrobiota",
  "Firmicutes" = "Bacillota",
  "Chloroflexi" = "Chloroflexota",
  "Halobacterota" = "Halobacterota",
  "Thermoplasmatota" = "Thermoplasmatota",
  "Patescibacteria" = "Patescibacteria",
  "Spirochaetota" = "Spirochaetota",
  "Armatimonadota" = "Armatimonadota",
  "Bdellovibrionota" = "Bdellovibrionota",
  "Gemmatimonadota" = "Gemmatimonadota",
  "Other" = "Other"
)

# plot
# ggplot(
#   ps_media_heat_df,
#   aes(
#     x = triplicate_id,
#     y = Phylum_ID,
#     fill = rel_abund
#   )
# ) +
#   geom_tile(color = "white", linewidth = 0.2) +
#   facet_wrap(~ SampleType + HRT, scales = "free_x", nrow = 1) +
#   scale_fill_gradient(
#     name = "Relative abundance",
#     low = "white",
#     high = "#1F77B4"
#   ) +
#   labs(
#     x = "Triplicate",
#     y = "Phylum"
#   ) +
#   theme_bw() +
#   theme(
#     panel.grid = element_blank(),
#     axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
#     axis.text = element_text(size = 12),
#     strip.text = element_text(size = 14),
#     axis.title = element_text(size = 14),
#     legend.title = element_text(size = 13),
#     legend.text = element_text(size = 11)
#   )

phyla_heatmap <- ggplot(
  ps_media_heat_df,
  aes(
    x = triplicate_id,
    y = Phylum_ID,
    fill = rel_abund_bin
  )
) +
  geom_tile(color = "white", linewidth = 0.2) +
  facet_wrap(~ SampleType + HRT, scales = "free_x", nrow = 1) +
  scale_fill_manual(
    name = "Relative abundance",
    values = heatmap_colors,
    drop = FALSE
  ) +
  scale_y_discrete(labels = new_phyla_taxa_names) +
  labs(
    x = "Triplicate",
    y = "Phylum"
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11)
  )

#ggsave("../raw_figs_301/phyla_317_heatmap.png", plot = phyla_heatmap, width = 14, height = 6, units = "in", dpi = 300)

# numeric means - Phyla relative abundance by sample type & HRT
ps_media_phyla_means <- ps_media_heat_df |>
  group_by(SampleType, HRT, Phylum_ID) |>
  summarise(mean_rel = mean(rel_abund, na.rm = TRUE),
            mean_rel_std = sd(rel_abund, na.rm = TRUE))
```

*CC & WC 317 se-asv Genera by triplicate ID*
```{r, fig.height=8, fig.width=12}
ps_media_rel_df <- ps_media |>
  psmelt() |>
  as_tibble() |>
  filter(!is.na(Genus), Genus != "NA")

# making a triplicate column
ps_media_rel_df$triplicate_id <- str_replace(ps_media_rel_df$SampleID, "_\\d+_(\\d+)", "_\\1")

ps_media_rel_df <- ps_media_rel_df |>
  group_by(triplicate_id) |>
  dplyr::mutate(relative_abund = Abundance / sum(Abundance, na.rm = TRUE)) |> 
  ungroup()

# assigning core vs other grouping <- all 317, not split by media-specific se-ASV
ps_media_rel_df$Genus_ID <- ifelse(
  ps_media_rel_df$OTU %in% se_asv_317, 
  as.character(ps_media_rel_df$Genus), 
  "Other"
)

ps_media_heat_df <- ps_media_rel_df |>
  group_by(
    SampleType,
    HRT,
    triplicate_id,
    Genus_ID
  ) |>
  dplyr::summarise(
    rel_abund = sum(relative_abund, na.rm = TRUE),
    .groups = "drop"
  )

# bin relative abundance for plot
ps_media_heat_df <- ps_media_heat_df %>%
  dplyr::mutate(
    rel_abund_bin = cut(
      rel_abund,
      breaks = c(0, 0.1, 0.2, 0.4, 0.6),
      include.lowest = TRUE,
      right = FALSE,
      labels = c(
        "0–0.1",
        "0.1–0.2",
        "0.2–0.4",
        "0.4–0.6"
      )
    )
  )

# bin color map
heatmap_colors <- c(
  "0–0.1"   = "#E0E9FF",
  "0.1–0.2" = "#ADC6FE",
  "0.2–0.4" = "#759FFD",
  "0.4–0.6" = "#2366FC"
)

ps_media_heat_df <- ps_media_heat_df %>%
  dplyr::mutate(
    rel_abund_bin = ifelse(
      rel_abund == 0,
      "0",
      as.character(rel_abund_bin)
    ),
    rel_abund_bin = factor(
      rel_abund_bin,
      levels = c("0", names(heatmap_colors))
    )
  )

heatmap_colors <- c(
  "0"       = "white",
  heatmap_colors
)


# optional: drop "other"
# ps_media_heat_df <- ps_media_heat_df |>
#   filter(Genus_ID != "Other")

# order genera top → bottom (reverse for heatmap aesthetics)

se_asv_genus_map <- c(
  "Xanthobacter", "Pseudorhodoplanes", "Azotobacter", "Burkholderia-Caballeronia-Paraburkholderia", "Cellulomonas", "Azospira", "Zoogloea", "Pseudomonas", "Plot4-2H12", 
  "Methanobacterium", "Mycobacterium", "Bradyrhizobium", "Nevskia", "Sphingomonas", "Geothrix", "Desulfovibrio", "Mesorhizobium", "Rhizomicrobium", "Caulobacter", 
  "Pleomorphomonas", "Coxiella", "Sediminibacterium", "Pajaroellobacter", "Undibacterium", "Edaphobaculum", "Pseudolabrys", "Kaistia", "Hyphomicrobium", "Rudaea", "Devosia", 
  "Aetherobacter", "Prosthecobacter", "Beijerinckia", "Puia", "Novosphingobium", "Dyella", "LD29", "Anaerolineaceae UCG-001", "Bacteroides", "Papillibacter", "Sphingobium", 
  "Gluconacetobacter", "Tahibacter", "Gemmata", "Rhodopila", "Ferruginibacter", "Thermincola", "WCHB1-32", "Limnohabitans",
  
  "Telmatospirillum", "Afipia", "Cytophaga", "Methylophilus", "Taibaiella", 
  "Methanocella", "Pir2 lineage", "Ellin516", "Bryobacter", 
  "Methanomassiliicoccus", "Methylocystis", "Bacillus", "Cupriavidus", 
  "Methanosarcina", "Chitinophaga", "Altererythrobacter", "Leptolinea", 
  "Terrimicrobium", "SN8", "Heliimonas", "Aquisphaera", "Paludisphaera", 
  "Fimbriiglobus", "Microvirgula", "Opitutus", "Chthoniobacter", 
  "Aquicella", "Ruminiclostridium", "Pseudoxanthobacter", "Terrimonas", 
  "Anaerospora", "Fibrisoma", "Thiobacillus", "Curvibacter", 
  "Pedomicrobium", "Solimonas", "Magnetospirillum", "Aminobacter", 
  "Desulfosporosinus", "Streptomyces", "Parachlamydia", 
  "Clostridium sensu stricto 1", "SM1A02", "NK4A214 group", "Ellin6055", 
  "Incertae Sedis", "Sphingoaurantiacus", "Singulisphaera", "Haliangium", 
  "Christensenellaceae R-7 group", "SH3-11", "Anaeromyxobacter", 
  "Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium", "Rhodoplanes", 
  "Clostridium sensu stricto 12", "Labrys", "Azospirillum", "Inquilinus", 
  "Steroidobacter", "Gordonia", "Thermomonas", "Nordella", "Rhodovastum", 
  "Psychrobacter", "Mucilaginibacter", "Galbitalea", "Methanospirillum", 
  "Acidibacter", "Sporobacter", "Williamsia", "Lacunisphaera", 
  "Defluviitaleaceae UCG-011", "Reyranella", "Bdellovibrio", "Anaerolinea", 
  "Spirosoma", "Gemmatimonas", "Methylobacterium-Methylorubrum", 
  "Sporomusa", "Brevundimonas", "Tundrisphaera", "Pseudobacteroides", 
  "Spirochaeta 2",
  
  "Other"
)

ps_media_heat_df$Genus_ID <- factor(
  ps_media_heat_df$Genus_ID,
  levels = rev(se_asv_genus_map))


# ordering triplicates
unique(ps_media_heat_df$triplicate_id)

triplicate_order <- c("4CC_0", "4CC_4", "4CC_7","4CC_11", "4CC_14", "4CC_18", "4CC_21",  
                      "CC_6", "CC_10", "CC_15", "CC_20",  
                      "CC_21", "CC_28", "CC_29", "CC_30", "CC_31", "CC_32", "CC_33", "CC_34", "CC_35", "CC_36", "CC_37", "CC_38", 
                      "4WC_0", "4WC_4", "4WC_8", "4WC_11", "4WC_14", "4WC_18", "4WC_21",  
                      "WC_6", "WC_10", "WC_15", "WC_20", 
                      "WC_21", "WC_24", "WC_27", "WC_28", "WC_29", "WC_30", "WC_31")

ps_media_heat_df$triplicate_id <- factor(ps_media_heat_df$triplicate_id, levels = triplicate_order)

genera_heatmap <- ggplot(
  ps_media_heat_df,
  aes(
    x = triplicate_id,
    y = Genus_ID,
    fill = rel_abund_bin
  )
) +
  geom_tile(color = "white", linewidth = 0.2) +
  facet_wrap(~ SampleType + HRT, scales = "free_x", nrow = 1) +
  scale_fill_manual(
    name = "Relative abundance",
    values = heatmap_colors,
    drop = FALSE
  ) +
  labs(
    x = "Triplicate",
    y = "Genus"
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
    axis.text.y = element_text(size = 6),
    strip.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11)
  )

#ggsave("../raw_figs_301/genera_317_heatmap.png", plot = genera_heatmap, width = 20, height = 13, units = "in", dpi = 300)
```












