---
title: "picrust2_prep"
output: html_document
date: "2026-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tutorial: https://shibalytics.com/teaching/16s/microbiome_analysis_part_v 

*Libraries*
```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ALDEx2")

if(!requireNamespace("BiocManager")){
  install.packages("BiocManager")
}
BiocManager::install("phyloseq")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("KEGGREST")

## ggpicrust2 version fix
# Remove the old version first
remove.packages("ggpicrust2")

# Reinstall from GitHub
devtools::install_github("cafferychen777/ggpicrust2")

# plotting issues
install.packages("ggprism")
```
```{r}
library(ggplot2)            #ggplot2_4.0.0 
library(tidyverse)          #tidyverse_2.0.0 
library(vegan)              #vegan_2.7-1 
library(data.table)         #data.table_1.17.8
library(reticulate)         #reticulate_1.43.0
library(picante)            #picante_1.8.2 
library(phyloseq)           #phyloseq_1.52.0 
library(ALDEx2)             #ALDEx2_1.40.0
library(kableExtra)         #kableExtra_1.4.0 
library(rstatix)            #rstatix_0.7.2 

library(ggpicrust2)         #ggpicrust2_2.5.2
library(ggprism)            #ggprism_1.0.7
library(patchwork)          #patchwork_1.3.2
library(readr)              #readr_2.1.5
library(MicrobiomeStat)     #MicrobiomeStat_1.2
library(KEGGREST)           #KEGGREST_1.48.1
library(GGally)             #GGally_2.4.0
library(tibble)             #tibble_3.3.0 
library(dplyr)              #dplyr_1.1.4
```

*PSO clean up functions*
```{r}
extract_sample_type <- function(id) {
  # Remove leading '4' if present
  id_no4 <- sub("^4", "", id)
  
  if (grepl("^CC", id_no4)) {
    return("corncob")
  } else if (grepl("^WC", id_no4)) {
    return("woodchip")
  } else if (grepl("^INF", id_no4)) {
    return("influent")
  } else {
    return(NA)
  }
}

extract_bioreactor_id <- function(id) {
  # If INF sample, reactor ID is "0"
  id_no4 <- sub("^4", "", id)
  
  if (grepl("^INF", id_no4)) {
    return("0")
  }
  
  matches <- regmatches(id_no4, regexec("^[A-Z]+_([0-9]+)_", id_no4))
  
  if (length(matches[[1]]) > 1) {
    return(matches[[1]][2])
  } else {
    return(NA)
  }
}

extract_HRT <- function(id) {
  if (grepl("^4", id)) {
    return(4)
  }
  
  if (grepl("^INF", id)) {
    num <- as.numeric(sub("^INF(\\d+)$", "\\1", id))
    if (!is.na(num)) {
      if (num <= 20) return(8)
      if (num >= 21) return(16)
    }
    return(NA)
  }
  
  day_str <- sub(".*_(\\d+)W?$", "\\1", id)  # pull numeric from end
  day <- as.numeric(day_str)
  
  if (!is.na(day)) {
    if (day >= 6 & day <= 20) {
      return(8)
    } else if (day >= 21) {
      return(16)
    }
  }
  
  return(NA) # if it doesn't match pattern
}

extract_sample_source <- function(id) {
  # Check if sample ID ends with "W"
  if (grepl("W$", id)) {
    return("water")
  }
  if (grepl("^INF", id)) {
    return("water")
  }
  else {
    return("media")
  }
}

extract_week <- function(id) {
  # 1. If sample starts with "INF", extract Week number after "INF"
  if (grepl("^INF", id)) {
    Week_str <- sub("^INF(\\d+)W?$", "\\1", id)
    Week <- as.numeric(Week_str)
    return(Week)
  }
  # 2. For other sample IDs, extract last number (before optional W at the end)
  Week_str <- sub(".*_(\\d+)W?$", "\\1", id)
  Week <- as.numeric(Week_str)
  if (!is.na(Week)) {
    return(Week)
  }
  return(NA) # return NA if no match
}
```

*Useful functions for generating PICRUSt2 input*
```{r}
#---- phyloseq_counts_to_df
# Converts phyloseq into a data frame which can be used to export the ASVs and samples with their counts to a file. This function also adds a new column with the ASV IDs to the front.
phyloseq_counts_to_df <- function(ps) {
  df <- as.data.frame(t(otu_table(ps)))
  df <- cbind("#OTU ID" = rownames(df), df)  # Cheeky way to add a comment to first line

  # Note: we use OTU ID above since at the moment this function is meanly used to pipe files into picrust, which needs the 'OTU ID'.
  return(df)
}

#---- write_counts_to_file
# Write the (raw) counts from a phyloseq object as a data matrix to file.
#
# INPUT  ps           : phyloseq object
#        file_name    : full file name (incl. path if required when not writing
#                       to working directory)
write_counts_to_file <- function(ps, filename) {
  df <- phyloseq_counts_to_df(ps)

  write.table(df,
              file=filename,
              sep = "\t",
              row.names = FALSE)
}

#---- write_seqs_to_file
# Write the sequences from a phyloseq object as a fasta file.
#
# INPUT  ps           : phyloseq object
#        filename    : full file name (incl. path if required when not writing
#                       to working directory)
write_seqs_to_file <- function(ps, filename) {
  df <- as.data.frame(refseq(ps))

  unlink(filename)  # make sure we delete before we concatenate below

  names <- rownames(df)
  for (i in 1:nrow(df)) {
    cat(paste0(">", names[i]), file=filename, sep="\n", append=TRUE)
    cat(df[i,1], file=filename, sep="\n", append=TRUE)
  }
  message(paste0("Wrote ", nrow(df), " ASV sequences to file ", filename))
}
```

*PSO read-in & sample data clean-up*
```{r}
# read in
ps_media_317 <- readRDS("../data/microbial/ps_media_317.rds") # 317 taxa, 114 samples

# samp df
ps_media_317_df <- sample_data(ps_media_317) |>
  dplyr::as_tibble() |>
  #dplyr::mutate and change string types, orders, etc. |>
  as.data.frame()

rownames(ps_media_317_df) = ps_media_317_df$SampleID
sample_data(ps_media_317) = data.frame(ps_media_317_df)
```

*Inspect and filter PSO*
```{r}
#--------------------------------------------
# Filter taxa (ASV) with low abundance (<2)
#--------------------------------------------
ps_media_317.p0 <- filter_taxa(ps_media_317, function (x) {sum(x > 0) >= 2}, prune=TRUE) # 317 taxa, 114 samples

#---------------------------------------------------
# Saving PICRUSt2 input files using functions above
#---------------------------------------------------
write_counts_to_file(ps_media_317, filename = "../data/picrust2/raw_counts_317.tsv")
write_seqs_to_file(ps_media_317, filename = "../data/picrust2/seqs_317.fna")
```

*################################################################################################################*
*################################################################################################################*
#-----------------------------------------------
### 6) PICRUSt2 - (Fig Prep/Separate Workflow)
# 317 media samples only, 16-h HRT
#-----------------------------------------------
- NOTE: must subset pso with 317 selected taxa first (code for this is under the section "4.5) P/A heatmaps")
- NOTE: pipeline is run in R terminal in HPC Interactive R session (a micromamba environment named "picrust2" was made beforehand)

```
#-------------------------------------
# Loading PICRUSt2 & running pipeline
#-------------------------------------

micromamba activate picrust2   # <- If starting session after installing 

#------------------------------------------------------------------------
# FULL PIPELINE (see DBRLC_PICRUSt2.Rmd for more details) - top 317 taxa
#------------------------------------------------------------------------

# NOTE: one ASV811 was lost
place_seqs.py -s ../../seqs_317.fna -o out_317.tre -p 2 \
  --intermediate intermediate_317/place_seqs \
  --chunk_size 1000
  
hsp.py -i 16S -t out_317.tre -o marker_predicted_and_nsti_317.tsv.gz -p 1 -n
hsp.py -i EC -t out_317.tre -o EC_predicted_317.tsv.gz -p 1
hsp.py -i KO -t out_317.tre -o KO_predicted_317.tsv.gz -p 1

# 2/316 ASVs were cut due NSTI cut-off of 2.0
metagenome_pipeline.py -i ../../raw_counts_317.tsv -m marker_predicted_and_nsti_317.tsv.gz -f EC_predicted_317.tsv.gz \
                       -o EC_metagenome_out_317 --strat_out

# 2/316 ASVs were cut due NSTI cut-off of 2.0
metagenome_pipeline.py \
  -i ../../raw_counts_317.tsv \
  -m marker_predicted_and_nsti_317.tsv.gz \
  -f KO_predicted_317.tsv.gz \
  -o KO_metagenome_out_317 \
  --strat_out

pathway_pipeline.py -i EC_metagenome_out_317/pred_metagenome_contrib.tsv.gz \
                    -o pathways_out_317 -p 1
                    
                    
add_descriptions.py -i EC_metagenome_out_317/pred_metagenome_unstrat.tsv.gz -m EC \
                    -o EC_metagenome_out_317/pred_metagenome_unstrat_descrip.tsv.gz

add_descriptions.py -i pathways_out_317/path_abun_unstrat.tsv.gz -m METACYC \
                    -o pathways_out_317/path_abun_unstrat_descrip.tsv.gz

add_descriptions.py -i KO_metagenome_out_317/pred_metagenome_unstrat.tsv.gz -m KO \
                    -o KO_metagenome_out_317/pred_metagenome_unstrat_descrip.tsv.gz

cp EC_metagenome_out_317/pred_metagenome_unstrat_descrip.tsv.gz  /work/adina/lilliang/DBRLC/picrust2/EC_317.tsv.gz
cp KO_metagenome_out_317/pred_metagenome_unstrat_descrip.tsv.gz  /work/adina/lilliang/DBRLC/picrust2/KO_317.tsv.gz
cp pathways_out_317/path_abun_unstrat_descrip.tsv.gz  /work/adina/lilliang/DBRLC/picrust2/path_317.tsv.gz

cp EC_metagenome_out_317/pred_metagenome_unstrat.tsv.gz  /work/adina/lilliang/DBRLC/picrust2/EC_num_317.tsv.gz
cp KO_metagenome_out_317/pred_metagenome_unstrat.tsv.gz  /work/adina/lilliang/DBRLC/picrust2/KO_num_317.tsv.gz
cp pathways_out_317/path_abun_unstrat.tsv.gz  /work/adina/lilliang/DBRLC/picrust2/path_num_317.tsv.gz
```



